---
title: "DFA Final"
author: "Megan J"
date: "July 28, 2020"
output: html_document
---


This document is organized in the following:
Chunks 1-2: Data import
Chunks 3-15: Graphs to visualize the data and patterns
Chunks 16-19: Running DFA for different trendline analyses, covariates, etc
Chunks 20-28: Graphs and analyses for the FOUR trendline analysis

The document is a cleaned and organized version of the DFA doc, removing those analyses for 3 trend lines and sample tests, and adding more DFA runs. Note that the looping DFA testing different R matrixes, 1-3 trends, and covaraites/no covariates is in DFA (not here)

```{r setup, include=FALSE}
library(MARSS)
library(lubridate)
library(dplyr)
library(ggplot2)
library(modelr)
require(gridExtra)
```

Below, the cleaned data is imported and prepared for analysis by transposing, converting dates, and Z-scoring. At the bottom, 3 years had additional cleaning that was done later to adjust for correlated data. These three years have a 'c'at the end of the data set year.

```{r Import Data}
#1. Read the data
#2. Summarize into by-day data for the backwater
#3. Change the date to date format
#4. Save the dates
#5. Transpose
#6. Set the column names as dates
#7. Z-score the data
#8. Save the discharge 

#-----2018 MC-------
MC.2018.raw<- read.csv("MC2018Day.csv") #1
MC.2018.raw$Date<- ymd(MC.2018.raw$Date) #3
MC.2018r<- as.matrix(MC.2018.raw[,c(2:6),]) 
date<- MC.2018.raw[,"Date"] #4
MC.2018.t<- t(MC.2018r) #5
colnames(MC.2018.t)<-date #6
MC.2018=(MC.2018.t-apply(MC.2018.t,1,mean,na.rm=TRUE))/ sqrt(apply(MC.2018.t,1,var,na.rm=TRUE)) #7
Discharge8 <- t(MC.2018.raw[,"Discharge",drop=FALSE]) #8


#-----2017 MC-------
MC.2017.raw<- read.csv("MC2017Day.csv") #1
MC.2017.raw$Date<- ymd(MC.2017.raw$Date) #3
MC.2017r<- as.matrix(MC.2017.raw[,c(2:6),])
date7<- MC.2017.raw[,"Date"] #4
MC.2017.t<- t(MC.2017r) #5
colnames(MC.2017.t)<-date7 #6
MC.2017=(MC.2017.t-apply(MC.2017.t,1,mean,na.rm=TRUE))/ sqrt(apply(MC.2017.t,1,var,na.rm=TRUE)) #7
Discharge7 <- t(MC.2017.raw[,"Discharge",drop=FALSE]) #8



#-----2016 MC-------
MC.2016.raw<- read.csv("MC2016Day.csv") #1
MC.2016.raw$Date<- ymd(MC.2016.raw$Date) #3
MC.2016r<- as.matrix(MC.2016.raw[,c(2:6),]) 
date6<- MC.2016.raw[,"Date"] #4
MC.2016.t<- t(MC.2016r) #5
colnames(MC.2016.t)<-date6 #6
MC.2016=(MC.2016.t-apply(MC.2016.t,1,mean,na.rm=TRUE))/ sqrt(apply(MC.2016.t,1,var,na.rm=TRUE)) #7
Discharge6 <- t(MC.2016.raw[,"Discharge",drop=FALSE]) #8



#-----2015 MC-------
MC.2015.raw<- read.csv("MC2015Day.csv") #1
MC.2015.raw$Date<- ymd(MC.2015.raw$Date) #3
MC.2015r<- as.matrix(MC.2015.raw[,c(2:6),]) 
date5<- MC.2015.raw[,"Date"] #4
MC.2015.t<- t(MC.2015r) #5
colnames(MC.2015.t)<-date5 #6
MC.2015=(MC.2015.t-apply(MC.2015.t,1,mean,na.rm=TRUE))/ sqrt(apply(MC.2015.t,1,var,na.rm=TRUE)) #7
Discharge5 <- t(MC.2015.raw[,"Discharge",drop=FALSE]) #8

#-----2018 SI-------
SI.2015.raw<- read.csv("SI2015B.csv") #1
SI.2015.raw<- SI.2015.raw %>% group_by(Date) %>% summarize(Turb= mean(Turb), CHLugL= mean(CHLugL), BGAugL= mean(BGAugL), 
                                                           FDOMqsu=mean(FDOMqsu), NO3_mgL=mean(NO3_mgL),
                                                           Discharge=mean(Discharge)) #2
SI.2015.raw$Date<- ymd(SI.2015.raw$Date) #3
SI.2015r<- as.matrix(SI.2015.raw[,c(2:6),])
dateS5<- ymd(SI.2015.raw$Date) #4
SI.2015.t<- t(SI.2015r) #5
colnames(SI.2015.t)<-dateS5 #6
SI.2015=(SI.2015.t-apply(SI.2015.t,1,mean,na.rm=TRUE))/ sqrt(apply(SI.2015.t,1,var,na.rm=TRUE)) #7
Discharge5s <- t(SI.2015.raw[,"Discharge",drop=FALSE]) #8



#-----2016 SI-------
SI.2016.raw<- read.csv("SI2016B.csv") #1
SI.2016.raw<- SI.2016.raw %>% group_by(Date) %>% summarize(Turb= mean(Turb), CHLugL= mean(CHLugL), BGAugL= mean(BGAugL), 
                                                           FDOMqsu=mean(FDOMqsu), NO3_mgL=mean(NO3_mgL),
                                                           Discharge=mean(Discharge)) #2
SI.2016.raw$Date<- ymd(SI.2016.raw$Date) #3
SI.2016r<- as.matrix(SI.2016.raw[,c(2:6),])
dateS6<- ymd(SI.2016.raw$Date) #4
SI.2016.t<- t(SI.2016r) #5
colnames(SI.2016.t)<-dateS6 #6
SI.2016=(SI.2016.t-apply(SI.2016.t,1,mean,na.rm=TRUE))/ sqrt(apply(SI.2016.t,1,var,na.rm=TRUE)) #7
Discharge6s <- t(SI.2016.raw[,"Discharge",drop=FALSE]) #8



#-----2015 SI-------
SI.2018.raw<- read.csv("SI2018B.csv") #1
SI.2018.raw<- SI.2018.raw %>% group_by(Date) %>% summarize(Turb= mean(Turb), CHLugL= mean(CHLugL), BGAugL= mean(BGAugL), 
                                                           FDOMqsu=mean(FDOMqsu), NO3_mgL=mean(NO3_mgL),
                                                           Discharge=mean(Discharge)) #2
SI.2018.raw$Date<- ymd(SI.2018.raw$Date) #3
SI.2018r<- as.matrix(SI.2018.raw[,c(2:6),])
dateS8<- ymd(SI.2018.raw$Date) #4
SI.2018.t<- t(SI.2018r) #5
colnames(SI.2018.t)<-dateS8 #6
SI.2018=(SI.2018.t-apply(SI.2018.t,1,mean,na.rm=TRUE))/ sqrt(apply(SI.2018.t,1,var,na.rm=TRUE)) #7
Discharge8s <- t(SI.2018.raw[,"Discharge",drop=FALSE]) #8



#--------CORRECTING FOR CORRELATED DATA--------------
SI.2015.rawc<- read.csv("2015SICalDay.csv") #1
SI.2015.rawc$Date<- ymd(SI.2015.rawc$Date) #3
SI.2015cr<- as.matrix(SI.2015.rawc[,c(2:6),])
dateSc5<- ymd(SI.2015.rawc$Date) #4
SI.2015.ct<- t(SI.2015cr) #5
colnames(SI.2015.ct)<-dateSc5 #6
SI.2015c=(SI.2015.ct-apply(SI.2015.ct,1,mean,na.rm=TRUE))/ sqrt(apply(SI.2015.ct,1,var,na.rm=TRUE)) #7
Discharge5cs <- t(SI.2015.rawc[,"Discharge",drop=FALSE]) #8

MC.2015.rawc<- read.csv("2015MCCalDay.csv") #1
MC.2015.rawc<-  MC.2015.rawc %>% filter(Date<'2015-10-01')
MC.2015.rawc<-  MC.2015.rawc %>% filter(!(Date=='2015-07-05'))

MC.2015.rawc$Date<- ymd(MC.2015.rawc$Date) #3
MC.2015cr<- as.matrix(MC.2015.rawc[,c(2:6),])
datec5<- ymd(MC.2015.rawc$Date) #4
MC.2015.ct<- t(MC.2015cr) #5
colnames(MC.2015.ct)<-datec5 #6
MC.2015c=(MC.2015.ct-apply(MC.2015.ct,1,mean,na.rm=TRUE))/ sqrt(apply(MC.2015.ct,1,var,na.rm=TRUE)) #7
Discharge5c <- t(MC.2015.rawc[,"Discharge",drop=FALSE]) #8

SI.2018.rawc<- read.csv("2018SICalDay.csv") #1
SI.2018.rawc$Date<- ymd(SI.2018.rawc$Date) #3
SI.2018cr<- as.matrix(SI.2018.rawc[,c(2:6),])
dateSc8<- ymd(SI.2018.rawc$Date) #4
SI.2018.ct<- t(SI.2018cr) #5
colnames(SI.2018.ct)<-dateSc8 #6
SI.2018c=(SI.2018.ct-apply(SI.2018.ct,1,mean,na.rm=TRUE))/ sqrt(apply(SI.2018.ct,1,var,na.rm=TRUE)) #7
Discharge8cs <- t(SI.2018.rawc[,"Discharge",drop=FALSE]) #8

MC.2015.rawc<-  MC.2015.rawc %>% filter(Date<'2015-10-01')
MC.2015.rawc<-  MC.2015.rawc %>% filter(!(Date=='2015-07-05'))
```

Raw data for seasonal plots, combined into AllData. The Type is the location and year, year and area (SI or MC) have also been added for easy separation/organizations

```{r Data for Seasonal Plots}

#Stroing data with the dates, April 13 is day 1
MC.2015.rawc<-  MC.2015.rawc %>% filter(Date<'2015-10-01')
MC.2015.rawc<-  MC.2015.rawc %>% filter(!(Date=='2015-07-05'))
MC.2016.raw<- MC.2016.raw %>% select(!month) %>% select(!day)
MC.2017.raw<- MC.2017.raw %>% select(!month) %>% select(!day)
AllData<- rbind(MC.2015.rawc %>% mutate( num= c(85:171)) %>% mutate(Type= c("MC2015") ) %>% mutate(Area= c("MC")) %>% mutate(Year= c("2015")), 
                   MC.2016.raw %>% mutate( num= c(1:204)) %>% mutate(Type= c("MC2016"))%>% mutate(Area= c("MC"))%>% mutate(Year= c("2016")),
                   MC.2017.raw %>% mutate( num= c(58:177)) %>% mutate(Type= c("MC2017"))%>% mutate(Area= c("MC"))%>% mutate(Year= c("2017")),
                   MC.2018.raw %>% mutate( num= c(50:171)) %>% mutate(Type= c("MC2018"))%>% mutate(Area= c("MC"))%>% mutate(Year= c("2018")),
                   SI.2015.rawc %>% mutate( num= c(50:186)) %>% mutate(Type= c("SI2015"))%>% mutate(Area= c("SI"))%>% mutate(Year= c("2015")), 
                   SI.2016.raw%>% mutate( num= c(1:185)) %>% mutate(Type= c("SI2016"))%>% mutate(Area= c("SI"))%>% mutate(Year= c("2016")),
                   SI.2018.rawc%>% mutate( num= c(40:178)) %>% mutate(Type= c("SI2018"))%>% mutate(Area= c("SI"))%>% mutate(Year= c("2018")))

#write.csv(AllData, "AllData.csv")

```

Boxplots showing the ranges for each variable. Each variable has its own boxplot, with the location and years all on that chart


```{r Boxplots}
#Showing ranges of each variable

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= NO3_mgL))+
  labs(x= "Location and Year", y= "Nitrate (mgL)", title= "Nitrate Levels")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= FDOMqsu))+
  labs(x= "Location and Year", y= "FDOM (qsu)", title= "FDOM Levels")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= CHLugL))+
  labs(x= "Location and Year", y= "Chlorophyll-a (ugL)", title= "Chlorophyll-a Levels")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= BGAugL))+
  labs(x= "Location and Year", y= "Blue-Green Algae (ugL)", title= "Blue-Green Algae Levels")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= Turb))+
  labs(x= "Location and Year", y= "Turbidity (NTU)", title= "Turbidity Levels")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))

ggplot(AllData) + geom_boxplot(mapping= aes(x=Type, y= Turb))+
  labs(x= "Location and Year", y= "Turbidity (NTU)", title= "Turbidity Levels Zoomed In")+
  scale_x_discrete(labels=c("Main Channel \n 2015", "Main Channel \n 2016", "Main Channel \n 2017", "Main Channel \n 2018", "Backwater \n 2015", "Backwater \n 2016", "Backwater \n 2018"))+ ylim(NA, 25)

ggplot(AllData) + geom_boxplot(mapping= aes(x=Year, y= Discharge))+
  labs(x= "Year", y= "Discharge (csu)", main= "Discharge Levels")+
  scale_x_discrete(labels=c(" 2015", "2016", "2017", "2018"))
  
```

Graphs of each variable graphed over the summer in 2017. Lines and annotations are added for the months, based on the 'num' value, which is the day

```{r Variables for 2017}
plotNMC7<-ggplot(data=AllDataMC7) + geom_line(mapping=aes(x=num, y=NO3_mgL), color= "#1C546B", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=.5, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Nitrate(mgL)")+
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Nitrate Levels")+
  
  theme(axis.title.y = element_text(size = 30))   +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotTMC7<-ggplot(data=AllDataMC7) + geom_line(mapping=aes(x=num, y=Turb), color= "#1C546B", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=.5, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity(NTU)")+
   theme(axis.title.y = element_text(size = 30)) + 
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Turbidity Levels")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotCMC7<-ggplot(data=AllDataMC7) + geom_line(mapping=aes(x=num, y=CHLugL), color= "#1C546B", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=.5, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Chlorophyll-a(ugL)")+
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Chlorophyll-a Levels")+
  
  theme(axis.title.y = element_text(size = 30))   +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotBMC7<-ggplot(data=AllDataMC7) + geom_line(mapping=aes(x=num, y=BGAugL), color= "#1C546B", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=.25, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Blue-Green Algae(ugL)")+
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Blue-Green Algae Levels")+
  
  theme(axis.title.y = element_text(size = 30))   +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotFMC7<-ggplot(data=AllDataMC7) + geom_line(mapping=aes(x=num, y=FDOMqsu), color= "#1C546B", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=62, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "FDOM(qsu)")+
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("FDOM Levels")+
  theme(axis.title.y = element_text(size = 30))   +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotNMC7
plotTMC7
plotCMC7
plotBMC7
plotFMC7
```

Each variable over time (4 lines for each year on the same plot)

```{r Variable for Years}

#MAIN Channel Nitrate for each year
AllDataMC<-AllData %>% filter(Area=='MC')
plotNMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=NO3_mgL, color= Type), size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Nitrate \n (mgL)")+
  theme(axis.title.x = element_blank()) +
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  #ggtitle("Nitrate Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#MAIN Channel Turb for each year
plotTMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=Turb/10, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=3, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity \n (10^1 NTU)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  scale_y_continuous(labels=scaleFUN) +
  
  theme(axis.title.x = element_blank()) +
   theme(legend.position= "none") +
  #ggtitle("Turbidity Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#MAIN Turb zoomed in
plotTZMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=Turb/100, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity \n (10^2 NTU)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  scale_y_continuous(labels=scaleFUN) +
  
  ylim(-.05,.30)+
  #ylim(-.05,.3)+
  #theme(axis.title.x = element_blank()) +
   theme(legend.position= "none") +
  #ggtitle("Turbidity Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())


#MAIN Channel BGA for each year
plotBMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=BGAugL, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Blue-Green \n Algae (ugL)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank() ) +
   theme(legend.position= "none") +
 # ggtitle("Blue-Green Algae Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#MAIN Channel CHL for each year
plotCMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=CHLugL/10, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Chlorophyll-a \n (10^1 ugL)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  scale_y_continuous(labels=scaleFUN) +
  
  theme(axis.title.x = element_blank()) +
   theme(legend.position= "none") +
  #ggtitle("Chlorophyll-a Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#MAIN Channel FDOM for each year
plotFMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=FDOMqsu/100, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=35, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "FDOM \n (10^2 qsu)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  
  theme(axis.title.x = element_blank()) +
  #ggtitle("FDOM Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotDMC<-ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=Discharge/100000, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13, 19, 50, 80, 111, 142, 171)+15), y=.1, label=c( "April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Discharge \n (10^5 cfu)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  #theme(axis.title.x = element_blank()) +
  #ggtitle("Discharge Levels Each Year")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())+
  ylim(-.1,NA)

grid.arrange(plotBMC, plotFMC, plotCMC,plotNMC,plotTZMC , plotDMC, nrow=3)
#Need to move dates, remove november

plotDMC

scaleFUN <- function(x) sprintf("%.1f", x)
```


Each variable over time (3 lines for each year on the same plot): BACKWATER

```{r Variable over BW years}
#BACK Nitrate each year
AllDataSI<-AllData %>% filter(Area=='SI')
plotNSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=NO3_mgL, color= Type), size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  
  labs(x= "Month", y= "Nitrate \n (mgL)")+
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  
  #ggtitle("Nitrate Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())


#BACK Turb each year
plotTSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=Turb, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity \n (NTU)") +
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  
  #ggtitle("Turbidity Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#BACK Turb Zoomed in each year
plotTZSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=Turb, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity \n (NTU)")+
  scale_y_continuous(labels=scaleFUN) +
  #theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  ylim(-.5, 9) +
  #ggtitle("Turbidity Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#BACK BGA each year
plotBSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=BGAugL, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  labs(x= "Month", y= "Blue-Green \n Algae (ugL)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  
  #ggtitle("Blue-Green Algae Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#BACK CHL each year
plotCSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=CHLugL/10, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
 # annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Chlorophyll-a \n (10^1 ugL)")+
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  #ggtitle("Chlorophyll-a Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#BACK FDOM
plotFSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=FDOMqsu/100, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  #annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=30, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  
  scale_y_continuous(labels=scaleFUN) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position= "none") +
  labs(x= "Month", y= "FDOM \n (10^2 qsu)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#009E73"), name= "Year", labels=c("2015", "2016", "2018")) +
  #ggtitle("FDOM Levels Each Year in the Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

plotDSI<-ggplot(data=AllDataSI) + geom_line(mapping=aes(x=num, y=Discharge/100000, color= Type), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Discharge \n (10^5 cfu)")+
  scale_color_manual(values= c("#999999", "#E69F00",  "#009E73"), name= "Year", labels=c("2015", "2016",  "2018")) +
 theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  #theme(axis.title.x = element_blank()) +
  #ggtitle("Discharge Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())+
  ylim(-.1,NA)
plotDSI
grid.arrange(plotBSI, plotFSI, plotCSI,plotNSI,plotTZSI , plotDSI, nrow=3)
```
Pairs of consituents graphed for each year and location. A function is created for each of the 10 pairs to see the pairing for each year and location


```{r Pairs Each Year}
AllDataMC5<-AllData %>% filter(Type=='MC2015')
AllDataMC6<-AllData %>% filter(Type=='MC2016')
AllDataMC7<-AllData %>% filter(Type=='MC2017')
AllDataMC8<-AllData %>% filter(Type=='MC2018')
AllDataSI5<-AllData %>% filter(Type=='SI2015')
AllDataSI6<-AllData %>% filter(Type=='SI2016')
AllDataSI8<-AllData %>% filter(Type=='SI2018')

PairGraphsBC<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=CHLugL), color=  "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=BGAugL*10), color= "#0072B2", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Chlorophyll-a",
    sec.axis = sec_axis(~./10, name="Blue-Green Algae (ugL)")
  ) + 
  ggtitle(paste("Chlorophyll-a and Blue-Green Algae in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color =  "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#0072B2", size=13)
  ) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())+ 
    labs(x="Month")
}
#"#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"

PairGraphsBC(AllDataMC5, '2015', 'Main Channel')
PairGraphsBC(AllDataMC6, '2016', 'Main Channel')
PairGraphsBC(AllDataMC7, '2017', 'Main Channel')
PairGraphsBC(AllDataMC8, '2018', 'Main Channel')
PairGraphsBC(AllDataSI5, '2015', 'Backwater')
PairGraphsBC(AllDataSI6, '2016', 'Backwater')
PairGraphsBC(AllDataSI8, '2018', 'Backwater')



#Scale is 50 for BW, 10 for MC
PairGraphsBF<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=FDOMqsu), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=BGAugL*50+100), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=95, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "FDOM (qsu)",
    sec.axis = sec_axis(~(.-100)/50, name="Blue-Green Algae (ugL)")
  ) + 
  ggtitle(paste("FDOM and Blue-Green Algae in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsBF(AllDataMC5, '2015', 'Main Channel')
PairGraphsBF(AllDataMC6, '2016', 'Main Channel')
PairGraphsBF(AllDataMC7, '2017', 'Main Channel')
PairGraphsBF(AllDataMC8, '2018', 'Main Channel')
PairGraphsBF(AllDataSI5, '2015', 'Backwater')
PairGraphsBF(AllDataSI6, '2016', 'Backwater')
PairGraphsBF(AllDataSI8, '2018', 'Backwater')

PairGraphsBN<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=NO3_mgL), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=BGAugL), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Nitrate (qsu)",
    sec.axis = sec_axis(~., name="Blue-Green Algae (ugL)")
  ) + 
  ggtitle(paste("Nitrate and Blue-Green Algae in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsBN(AllDataMC5, '2015', 'Main Channel')
PairGraphsBN(AllDataMC6, '2016', 'Main Channel')
PairGraphsBN(AllDataMC7, '2017', 'Main Channel')
PairGraphsBN(AllDataMC8, '2018', 'Main Channel')
PairGraphsBN(AllDataSI5, '2015', 'Backwater')
PairGraphsBN(AllDataSI6, '2016', 'Backwater')
PairGraphsBN(AllDataSI8, '2018', 'Backwater')

PairGraphsBT<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Turb), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=BGAugL+5), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Turbidity (NTU)",
    sec.axis = sec_axis(~.-5, name="Blue-Green Algae (ugL)")
  ) + 
    ylim(NA, 15)+
  ggtitle(paste("Turbidity and Blue-Green Algae in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsBT(AllDataMC5, '2015', 'Main Channel')
PairGraphsBT(AllDataMC6, '2016', 'Main Channel')
PairGraphsBT(AllDataMC7, '2017', 'Main Channel')
PairGraphsBT(AllDataMC8, '2018', 'Main Channel')
PairGraphsBT(AllDataSI5, '2015', 'Backwater')
PairGraphsBT(AllDataSI6, '2016', 'Backwater')
PairGraphsBT(AllDataSI8, '2018', 'Backwater')


PairGraphsCF<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=CHLugL), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(FDOMqsu-100)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Chlorophyll (ugL)",
    sec.axis = sec_axis(~.+100, name="FDOM (qsu)")
  ) + 
    
  ggtitle(paste("FDOM and Chlorophyll-a in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsCF(AllDataMC5, '2015', 'Main Channel')
PairGraphsCF(AllDataMC6, '2016', 'Main Channel')
PairGraphsCF(AllDataMC7, '2017', 'Main Channel')
PairGraphsCF(AllDataMC8, '2018', 'Main Channel')
PairGraphsCF(AllDataSI5, '2015', 'Backwater')
PairGraphsCF(AllDataSI6, '2016', 'Backwater')
PairGraphsCF(AllDataSI8, '2018', 'Backwater')


PairGraphsCN<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=CHLugL), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(NO3_mgL*10)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Chlorophyll-a (ugL)",
    sec.axis = sec_axis(~./10, name="Nitrate (mgL)")
  ) + 
    
  ggtitle(paste("Chlorophyll-a and Nitrate in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsCN(AllDataMC5, '2015', 'Main Channel')
PairGraphsCN(AllDataMC6, '2016', 'Main Channel')
PairGraphsCN(AllDataMC7, '2017', 'Main Channel')
PairGraphsCN(AllDataMC8, '2018', 'Main Channel')
PairGraphsCN(AllDataSI5, '2015', 'Backwater')
PairGraphsCN(AllDataSI6, '2016', 'Backwater')
PairGraphsCN(AllDataSI8, '2018', 'Backwater')


PairGraphsCT<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=CHLugL), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(Turb*4)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Chlorophyll-a (ugL)",
    sec.axis = sec_axis(~./4, name="Turbidity (NTU)")
  ) + 
    #ylim(NA,100)+ #For BW 2018
  ggtitle(paste("Chlorophyll-a and Turbidty in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsCT(AllDataMC5, '2015', 'Main Channel')
PairGraphsCT(AllDataMC6, '2016', 'Main Channel')
PairGraphsCT(AllDataMC7, '2017', 'Main Channel')
PairGraphsCT(AllDataMC8, '2018', 'Main Channel')
PairGraphsCT(AllDataSI5, '2015', 'Backwater')
PairGraphsCT(AllDataSI6, '2016', 'Backwater')
PairGraphsCT(AllDataSI8, '2018', 'Backwater')


PairGraphsFN<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=FDOMqsu), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=((NO3_mgL*5)+100)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=100, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "FDOM (qsu)",
    sec.axis = sec_axis(~(.-100)/5, name="Nitrate (mgL)")
  ) + 
    
  ggtitle(paste("FDOM and Nitrate in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsFN(AllDataMC5, '2015', 'Main Channel')
PairGraphsFN(AllDataMC6, '2016', 'Main Channel')
PairGraphsFN(AllDataMC7, '2017', 'Main Channel')
PairGraphsFN(AllDataMC8, '2018', 'Main Channel')
PairGraphsFN(AllDataSI5, '2015', 'Backwater')
PairGraphsFN(AllDataSI6, '2016', 'Backwater')
PairGraphsFN(AllDataSI8, '2018', 'Backwater')


PairGraphsFT<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=FDOMqsu), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(Turb*10)+20), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=100, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "FDOM (qsu)",
    sec.axis = sec_axis(~(.-20)/10, name="Turbidity (NTU)")
  ) + 
    #ylim(NA, 250)+ #For BW 2018
  ggtitle(paste("FDOM and Turbidity in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsFT(AllDataMC5, '2015', 'Main Channel')
PairGraphsFT(AllDataMC6, '2016', 'Main Channel')
PairGraphsFT(AllDataMC7, '2017', 'Main Channel')
PairGraphsFT(AllDataMC8, '2018', 'Main Channel')
PairGraphsFT(AllDataSI5, '2015', 'Backwater')
PairGraphsFT(AllDataSI6, '2016', 'Backwater')
PairGraphsFT(AllDataSI8, '2018', 'Backwater')


PairGraphsNT<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=NO3_mgL), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(Turb/4)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Nitrate (mgL)",
    sec.axis = sec_axis(~./4, name="Turbidity (NTU)")
  ) + 
  
  ggtitle(paste("Nitrate and Turbidity in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsNT(AllDataMC5, '2015', 'Main Channel')
PairGraphsNT(AllDataMC6, '2016', 'Main Channel')
PairGraphsNT(AllDataMC7, '2017', 'Main Channel')
PairGraphsNT(AllDataMC8, '2018', 'Main Channel')
PairGraphsNT(AllDataSI5, '2015', 'Backwater')
PairGraphsNT(AllDataSI6, '2016', 'Backwater')
PairGraphsNT(AllDataSI8, '2018', 'Backwater')
```

Finding the correlations between each pairing for each year and locations

```{r Pairs correlations}
#Blue-green algae, Chl
cor(AllDataMC5$CHLugL, AllDataMC5$BGAugL)
cor(AllDataMC6$CHLugL, AllDataMC6$BGAugL)
cor(AllDataMC7$CHLugL, AllDataMC7$BGAugL)
cor(AllDataMC8$CHLugL, AllDataMC8$BGAugL)
cor(AllDataSI5$CHLugL, AllDataSI5$BGAugL)
cor(AllDataSI6$CHLugL, AllDataSI6$BGAugL)
cor(AllDataSI8$CHLugL, AllDataSI8$BGAugL)

#Blue-green algae, FDOM
cor(AllDataMC5$FDOMqsu, AllDataMC5$BGAugL)
cor(AllDataMC6$FDOMqsu, AllDataMC6$BGAugL)
cor(AllDataMC7$FDOMqsu, AllDataMC7$BGAugL)
cor(AllDataMC8$FDOMqsu, AllDataMC8$BGAugL)
cor(AllDataSI5$FDOMqsu, AllDataSI5$BGAugL)
cor(AllDataSI6$FDOMqsu, AllDataSI6$BGAugL)
cor(AllDataSI8$FDOMqsu, AllDataSI8$BGAugL)

#Blue-green algae, nitrate
cor(AllDataMC5$NO3_mgL, AllDataMC5$BGAugL)
cor(AllDataMC6$NO3_mgL, AllDataMC6$BGAugL)
cor(AllDataMC7$NO3_mgL, AllDataMC7$BGAugL)
cor(AllDataMC8$NO3_mgL, AllDataMC8$BGAugL)
cor(AllDataSI5$NO3_mgL, AllDataSI5$BGAugL)
cor(AllDataSI6$NO3_mgL, AllDataSI6$BGAugL)
cor(AllDataSI8$NO3_mgL, AllDataSI8$BGAugL)

#Blue-green algae, turbidity
cor(AllDataMC5$Turb, AllDataMC5$BGAugL)
cor(AllDataMC6$Turb, AllDataMC6$BGAugL)
cor(AllDataMC7$Turb, AllDataMC7$BGAugL)
cor(AllDataMC8$Turb, AllDataMC8$BGAugL)
cor(AllDataSI5$Turb, AllDataSI5$BGAugL)
cor(AllDataSI6$Turb, AllDataSI6$BGAugL)
cor(AllDataSI8$Turb, AllDataSI8$BGAugL)

#Blue-green algae, discharge
cor(AllDataMC5$Discharge, AllDataMC5$BGAugL)
cor(AllDataMC6$Discharge, AllDataMC6$BGAugL)
cor(AllDataMC7$Discharge, AllDataMC7$BGAugL)
cor(AllDataMC8$Discharge, AllDataMC8$BGAugL)
cor(AllDataSI5$Discharge, AllDataSI5$BGAugL)
cor(AllDataSI6$Discharge, AllDataSI6$BGAugL)
cor(AllDataSI8$Discharge, AllDataSI8$BGAugL)

#CHL, FDOM
cor(AllDataMC5$CHLugL, AllDataMC5$FDOMqsu)
cor(AllDataMC6$CHLugL, AllDataMC6$FDOMqsu)
cor(AllDataMC7$CHLugL, AllDataMC7$FDOMqsu)
cor(AllDataMC8$CHLugL, AllDataMC8$FDOMqsu)
cor(AllDataSI5$CHLugL, AllDataSI5$FDOMqsu)
cor(AllDataSI6$CHLugL, AllDataSI6$FDOMqsu)
cor(AllDataSI8$CHLugL, AllDataSI8$FDOMqsu)

#CHL, Nitrate
cor(AllDataMC5$CHLugL, AllDataMC5$NO3_mgL)
cor(AllDataMC6$CHLugL, AllDataMC6$NO3_mgL)
cor(AllDataMC7$CHLugL, AllDataMC7$NO3_mgL)
cor(AllDataMC8$CHLugL, AllDataMC8$NO3_mgL)
cor(AllDataSI5$CHLugL, AllDataSI5$NO3_mgL)
cor(AllDataSI6$CHLugL, AllDataSI6$NO3_mgL)
cor(AllDataSI8$CHLugL, AllDataSI8$NO3_mgL)

#CHL, Turb
cor(AllDataMC5$CHLugL, AllDataMC5$Turb)
cor(AllDataMC6$CHLugL, AllDataMC6$Turb)
cor(AllDataMC7$CHLugL, AllDataMC7$Turb)
cor(AllDataMC8$CHLugL, AllDataMC8$Turb)
cor(AllDataSI5$CHLugL, AllDataSI5$Turb)
cor(AllDataSI6$CHLugL, AllDataSI6$Turb)
cor(AllDataSI8$CHLugL, AllDataSI8$Turb)

#CHL, Discharge
cor(AllDataMC5$CHLugL, AllDataMC5$Discharge)
cor(AllDataMC6$CHLugL, AllDataMC6$Discharge)
cor(AllDataMC7$CHLugL, AllDataMC7$Discharge)
cor(AllDataMC8$CHLugL, AllDataMC8$Discharge)
cor(AllDataSI5$CHLugL, AllDataSI5$Discharge)
cor(AllDataSI6$CHLugL, AllDataSI6$Discharge)
cor(AllDataSI8$CHLugL, AllDataSI8$Discharge)

#FDOM, Nitrate


#FDOM, turb

#FDOM, Discharge

#Nitrate, turb

#Nitrate, discharge

#Turb, discharge


```


Graphs of the main channel vs backwater. Each consitutent has a function to graph the mc vs bw for each year
```{r MC vs BW}

#Main channel vs backwater for each variable, for each year
AllData5<-AllData %>% filter(Year=='2015')
AllData6<-AllData %>% filter(Year=='2016')
AllData8<-AllData %>% filter(Year=='2018')

LocationN03<- function(datax, yr){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=NO3_mgL, color = Area), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Nitrate (mgL)")+
  ggtitle(paste("Nitrate in", yr,  "in the Main channel and Backwater", sep=" "))+
    scale_color_manual(values= c("#E69F00","#56B4E9" ), name= "Location", labels=c("Main Channel", "Backwater")) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}
LocationN03(AllData5, "2015")
LocationN03(AllData6, "2016")
LocationN03(AllData8, "2018")


LocationFDOM<- function(datax, yr){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=FDOMqsu, color = Area), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=90, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "FDOM (qsu)")+
  ggtitle(paste("FDOM in", yr,  "in the Main channel and Backwater", sep=" "))+
    scale_color_manual(values= c("#E69F00","#56B4E9" ), name= "Location", labels=c("Main Channel", "Backwater")) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}
LocationFDOM(AllData5, "2015")
LocationFDOM(AllData6, "2016")
LocationFDOM(AllData8, "2018")


LocationCHL<- function(datax, yr){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=CHLugL, color = Area), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Chlorophyll-a (ugL)")+
  ggtitle(paste("Chlorophyll-a Levels for Each Location in", yr, sep=" "))+
    scale_color_manual(values= c("#E69F00","#56B4E9" ), name= "Location", labels=c("Main Channel", "Backwater")) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}
LocationCHL(AllData5, "2015")
LocationCHL(AllData6, "2016")
LocationCHL(AllData8, "2018")


LocationBGA<- function(datax, yr){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=BGAugL, color = Area), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Blue-Green Algae (ugL)")+
  ggtitle(paste("Blue-Green Algae in", yr,  "in the Main channel and Backwater", sep=" "))+
    scale_color_manual(values= c("#E69F00","#56B4E9" ), name= "Location", labels=c("Main Channel", "Backwater")) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}
LocationBGA(AllData5, "2015")
LocationBGA(AllData6, "2016")
LocationBGA(AllData8, "2018")



LocationTurb<- function(datax, yr){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Turb, color = Area), size=1) + 
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=0, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Turbidity (NTU)")+
  ggtitle(paste("Turbidity in", yr,  "in the Main channel and Backwater", sep=" "))+
    scale_color_manual(values= c("#E69F00","#56B4E9" ), name= "Location", labels=c("Main Channel", "Backwater")) +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}
LocationTurb(AllData5, "2015")
LocationTurb(AllData6, "2016")
LocationTurb(AllData8, "2018")


```
Correlations between mc and bw for each year and each constituent
```{r MC BW Correlations}
AllDataSI5Short<- AllDataSI5 %>% filter(!(Date<'2015-07-06'), !(Date>'2015-09-30'))
AllDataMC6Short<- AllDataMC6 %>% filter( !(Date>'2016-10-14'))
AllDataSI8Short<- AllDataSI8 %>% filter(!(Date<'2018-06-01'), !(Date>'2018-09-30'))

cor(AllDataMC5$BGAugL, AllDataSI5Short$BGAugL)
cor(AllDataMC6Short$BGAugL, AllDataSI6$BGAugL)
cor(AllDataMC8$BGAugL, AllDataSI8Short$BGAugL) #Not strong

cor(AllDataMC5$CHLugL, AllDataSI5Short$CHLugL)
cor(AllDataMC6Short$CHLugL, AllDataSI6$CHLugL)
cor(AllDataMC8$CHLugL, AllDataSI8Short$CHLugL) #Not strong

cor(AllDataMC5$FDOMqsu, AllDataSI5Short$FDOMqsu) #Not strong
cor(AllDataMC6Short$FDOMqsu, AllDataSI6$FDOMqsu)#Strong
cor(AllDataMC8$FDOMqsu, AllDataSI8Short$FDOMqsu) #Strong

cor(AllDataMC5$NO3_mgL, AllDataSI5Short$NO3_mgL) 
cor(AllDataMC6Short$NO3_mgL, AllDataSI6$NO3_mgL)
cor(AllDataMC8$NO3_mgL, AllDataSI8Short$NO3_mgL)#Strong

cor(AllDataMC5$Turb, AllDataSI5Short$Turb)
cor(AllDataMC6Short$Turb, AllDataSI6$Turb)
cor(AllDataMC8$Turb, AllDataSI8Short$Turb) #Not strong
```

Plots of discharge against each variable. Each variable has a function to graph discharge vs the variable for each year and location

```{r Discharge plots}

#Discharge vs each constituent for each year/location
PairGraphsDB<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Discharge), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(BGAugL*10000)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Discharge (csu)",
    sec.axis = sec_axis(~./10000, name="Blue-Green Algae (ugL)")
  ) + 
  
  ggtitle(paste("Blue-Green Algae and Discharge in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsDB(AllDataMC5, '2015', 'Main Channel')
PairGraphsDB(AllDataMC6, '2016', 'Main Channel')
PairGraphsDB(AllDataMC7, '2017', 'Main Channel')
PairGraphsDB(AllDataMC8, '2018', 'Main Channel')
PairGraphsDB(AllDataSI5, '2015', 'Backwater')
PairGraphsDB(AllDataSI6, '2016', 'Backwater')
PairGraphsDB(AllDataSI8, '2018', 'Backwater')

PairGraphsDC<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Discharge), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(CHLugL*500)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Discharge (csu)",
    sec.axis = sec_axis(~./500, name="Chlorophyll-a (ugL)")
  ) + 
  
  ggtitle(paste("Chlorophyll-a and Discharge in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsDC(AllDataMC5, '2015', 'Main Channel')
PairGraphsDC(AllDataMC6, '2016', 'Main Channel')
PairGraphsDC(AllDataMC7, '2017', 'Main Channel')
PairGraphsDC(AllDataMC8, '2018', 'Main Channel')
PairGraphsDC(AllDataSI5, '2015', 'Backwater')
PairGraphsDC(AllDataSI6, '2016', 'Backwater')
PairGraphsDC(AllDataSI8, '2018', 'Backwater')


PairGraphsDF<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Discharge), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(FDOMqsu*250)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Discharge (csu)",
    sec.axis = sec_axis(~./250, name="FDOM (qsu)")
  ) + 
  
  ggtitle(paste("FDOM and Discharge in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsDF(AllDataMC5, '2015', 'Main Channel')
PairGraphsDF(AllDataMC6, '2016', 'Main Channel')
PairGraphsDF(AllDataMC7, '2017', 'Main Channel')
PairGraphsDF(AllDataMC8, '2018', 'Main Channel')
PairGraphsDF(AllDataSI5, '2015', 'Backwater')
PairGraphsDF(AllDataSI6, '2016', 'Backwater')
PairGraphsDF(AllDataSI8, '2018', 'Backwater')


PairGraphsDN<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Discharge), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(NO3_mgL*10000)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Discharge (csu)",
    sec.axis = sec_axis(~./10000, name="Nitrate (mgL)")
  ) + 
  
  ggtitle(paste("Nitrate and Discharge in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsDN(AllDataMC5, '2015', 'Main Channel')
PairGraphsDN(AllDataMC6, '2016', 'Main Channel')
PairGraphsDN(AllDataMC7, '2017', 'Main Channel')
PairGraphsDN(AllDataMC8, '2018', 'Main Channel')
PairGraphsDN(AllDataSI5, '2015', 'Backwater')
PairGraphsDN(AllDataSI6, '2016', 'Backwater')
PairGraphsDN(AllDataSI8, '2018', 'Backwater')


PairGraphsDT<- function(datax, yr, location){
  ggplot(data=datax) + geom_line(mapping=aes(x=num, y=Discharge), color= "#56B4E9", size=1) + 
   geom_line(mapping=aes(x=num, y=(Turb*5000)), color= "#E69F00", size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  scale_y_continuous(
    name = "Discharge (csu)",
    sec.axis = sec_axis(~./5000, name="Turbidity (NTU)")
  ) + 
  
  ggtitle(paste("Turbidity and Discharge in", yr,  "in the", location, sep=" "))+
  theme(
    axis.title.y = element_text(color = "#56B4E9", size=13),
    axis.title.y.right = element_text(color = "#E69F00", size=13)
  ) +
    labs(x="Month") +
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
}

PairGraphsDF(AllDataMC5, '2015', 'Main Channel')
PairGraphsDF(AllDataMC6, '2016', 'Main Channel')
PairGraphsDF(AllDataMC7, '2017', 'Main Channel')
PairGraphsDF(AllDataMC8, '2018', 'Main Channel')
PairGraphsDF(AllDataSI5, '2015', 'Backwater')
PairGraphsDF(AllDataSI6, '2016', 'Backwater')
PairGraphsDF(AllDataSI8, '2018', 'Backwater')


ggplot(data=AllDataMC) + geom_line(mapping=aes(x=num, y=Discharge, color= Type), size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=10000, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Discharge (csu)")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  
  ggtitle("Discharge Levels Each Year in the Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

```

Dot plot of each variable against another variable. Plots include data from all years for the main channel and a trendline

```{r Relationship of Variables MC}
#Dotplot with relationships of pairs

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=BGAugL, y=CHLugL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Chlorophyll-a (ugL)", title= "Chlorophyll-a vs Blue-Green Algae in the Main Channel")+
  geom_smooth(aes(x=BGAugL, y=CHLugL, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=BGAugL, y=FDOMqsu, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="FDOM (qsu)", title= "FDOM vs Blue-Green Algae in the Main Channel")+
  geom_smooth(aes(x=BGAugL, y=FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=BGAugL, y=NO3_mgL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Blue-Green Algae in the Main Channel")+
  geom_smooth(aes(x=BGAugL, y=NO3_mgL, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=BGAugL, y=Turb, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Blue-Green Algae in the Main Channel")+
  geom_smooth(aes(x=BGAugL, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA,30)

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=BGAugL, y=Discharge, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Discharge (csu)", title= "Discharge vs Blue-Green Algae in the Main Channel")+
  geom_smooth(aes(x=BGAugL, y=Discharge, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=CHLugL, y=FDOMqsu, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="FDOM (qsu)", title= "FDOM vs Chlorophyll-a in the Main Channel")+
  geom_smooth(aes(x=CHLugL, FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=CHLugL, y=NO3_mgL, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Chlorophyll-a in the Main Channel")+
  geom_smooth(aes(x=CHLugL, NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=CHLugL, y=Turb, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Chlorophyll-a in the Main Channel")+
  geom_smooth(aes(x=CHLugL, Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 35)

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=CHLugL, y=Discharge, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Discharge (csu)", title= "Discharge vs Chlorophyll-a in the Main Channel")+
  geom_smooth(aes(x=CHLugL, Discharge, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=FDOMqsu, y=NO3_mgL, color=Year))+
  labs(x="FDOM (qsu)", y= "Nitrate (mgL)", title= "Nitrate vs FDOM in the Main Channel")+
  geom_smooth(aes(x=FDOMqsu, y=NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=FDOMqsu, y=Turb, color=Year))+
  labs(x="FDOM (qsu)", y= "Turbidity (NTU)", title= "Turbidity vs FDOM in the Main Channel")+
  geom_smooth(aes(x=FDOMqsu, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 30)

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=FDOMqsu, y=Discharge, color=Year))+
  labs(x="FDOM (qsu)", y= "Discharge (csu)", title= "Discharge vs FDOM in the Main Channel")+
  geom_smooth(aes(x=FDOMqsu, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=NO3_mgL, y=Turb, color=Year))+
  labs(x="Nitrate (mgL)", y= "Turbidity (NTU)", title= "Turbidity vs Nitrate in the Main Channel")+
  geom_smooth(aes(x=NO3_mgL, y=Turb, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 25)

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=NO3_mgL, y=Discharge, color=Year))+
  labs(x="Nitrate (mgL)", y= "Discharge (csu)", title= "Discharge vs Nitrate in the Main Channel")+
  geom_smooth(aes(x=NO3_mgL, y=Discharge, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataMC) +
  geom_point(mapping=aes(x=Turb, y=Discharge, color=Year))+
  labs(x="Turbidity (NTU)", y= "Discharge (csu)", title= "Discharge vs Turbidity in the Main Channel")+
  geom_smooth(aes(x=Turb, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  xlim(NA, 30)
```
Dot plot of each variable against another variable. Plots include data from all years for the backwater and a trendline
```{r Relationship of Variables BW}
#Dotplots of pairs of variables 

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=BGAugL, y=CHLugL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Chlorophyll-a (ugL)", title= "Chlorophyll-a vs Blue-Green Algae in the Backwater")+
  geom_smooth(aes(x=BGAugL, y=CHLugL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=BGAugL, y=FDOMqsu, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="FDOM (qsu)", title= "FDOM vs Blue-Green Algae in the Backwater")+
  geom_smooth(aes(x=BGAugL, y=FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=BGAugL, y=NO3_mgL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Blue-Green Algae in the Backwater")+
  geom_smooth(aes(x=BGAugL, y=NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=BGAugL, y=Turb, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Blue-Green Algae in the Backwater")+
  geom_smooth(aes(x=BGAugL, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA,30)

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=BGAugL, y=Discharge, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Discharge (csu)", title= "Discharge vs Blue-Green Algae in the Backwater")+
  geom_smooth(aes(x=BGAugL, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=CHLugL, y=FDOMqsu, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="FDOM (qsu)", title= "FDOM vs Chlorophyll-a in the Backwater")+
  geom_smooth(aes(x=CHLugL, FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=CHLugL, y=NO3_mgL, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Chlorophyll-a in the Backwater")+
  geom_smooth(aes(x=CHLugL, NO3_mgL, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=CHLugL, y=Turb, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Chlorophyll-a in the Backwater")+
  geom_smooth(aes(x=CHLugL, Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 35)

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=CHLugL, y=Discharge, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Discharge (csu)", title= "Discharge vs Chlorophyll-a in the Backwater")+
  geom_smooth(aes(x=CHLugL, Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=FDOMqsu, y=NO3_mgL, color=Year))+
  labs(x="FDOM (qsu)", y= "Nitrate (mgL)", title= "Nitrate vs FDOM in the Backwater")+
  geom_smooth(aes(x=FDOMqsu, y=NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=FDOMqsu, y=Turb, color=Year))+
  labs(x="FDOM (qsu)", y= "Turbidity (NTU)", title= "Turbidity vs FDOM in the Backwater")+
  geom_smooth(aes(x=FDOMqsu, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 30)

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=FDOMqsu, y=Discharge, color=Year))+
  labs(x="FDOM (qsu)", y= "Discharge (csu)", title= "Discharge vs FDOM in the Backwater")+
  geom_smooth(aes(x=FDOMqsu, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=NO3_mgL, y=Turb, color=Year))+
  labs(x="Nitrate (mgL)", y= "Turbidity (NTU)", title= "Turbidity vs Nitrate in the Backwater")+
  geom_smooth(aes(x=NO3_mgL, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 25)

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=NO3_mgL, y=Discharge, color=Year))+
  labs(x="Nitrate (mgL)", y= "Discharge (csu)", title= "Discharge vs Nitrate in the Backwater")+
  geom_smooth(aes(x=NO3_mgL, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllDataSI) +
  geom_point(mapping=aes(x=Discharge, y=Turb, color=Year))+
  labs(x="Turbidity (NTU)", y= "Discharge (csu)", title= "Discharge vs Turbidity in the Backwater")+
  geom_smooth(aes(x=Discharge, y=Turb, color=Year),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 10)
```
Dot plot of each variable against another variable. Plots include data from all years for the main channel and backwater and a trendline
```{r Relationship of Variables Both}
#Dotplots for both

ggplot(data=AllData) +
  geom_point(mapping=aes(x=BGAugL, y=CHLugL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Chlorophyll-a (ugL)", title= "Chlorophyll-a vs Blue-Green Algae for All Locations")+
  geom_smooth(aes(x=BGAugL, y=CHLugL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=BGAugL, y=FDOMqsu, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="FDOM (qsu)", title= "FDOM vs Blue-Green Algae for All Locations")+
  geom_smooth(aes(x=BGAugL, y=FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=BGAugL, y=NO3_mgL, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Blue-Green Algae for All Locations")+
  geom_smooth(aes(x=BGAugL, y=NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=BGAugL, y=Turb, color=Year))+
  labs(x="Blue-Green Algae (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Blue-Green Algae for All Locations")+
  geom_smooth(aes(x=BGAugL, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA,30)

ggplot(data=AllData) +
  geom_point(mapping=aes(x=Discharge, y=BGAugL, color=Year))+
  labs(x= "Discharge (csu)", y= "Blue-Green Algae (ugL)", title= "Discharge vs Blue-Green Algae for All Locations")+
  geom_smooth(aes(x=Discharge, y=BGAugL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#0072B2", "#56B4E9", "#009E73", "#D55E00"))+
  annotate(geom="label", x=100000, y=4, label="r = -0.41")

ggplot(data=AllData) +
  geom_point(mapping=aes(x=CHLugL, y=FDOMqsu, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="FDOM (qsu)", title= "FDOM vs Chlorophyll-a for All Locations")+
  geom_smooth(aes(x=CHLugL, FDOMqsu),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=CHLugL, y=NO3_mgL, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Nitrate (mgL)", title= "Nitrate vs Chlorophyll-a for All Locations")+
  geom_smooth(aes(x=CHLugL, NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=CHLugL, y=Turb, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Turbidity (NTU)", title= "Turbidity vs Chlorophyll-a for All Locations")+
  geom_smooth(aes(x=CHLugL, Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 35)

ggplot(data=AllData) +
  geom_point(mapping=aes(x=CHLugL, y=Discharge, color=Year))+
  labs(x="Chlorophyll-a (ugL)", y="Discharge (csu)", title= "Discharge vs Chlorophyll-a for All Locations")+
  geom_smooth(aes(x=CHLugL, Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=FDOMqsu, y=NO3_mgL, color=Year))+
  labs(x="FDOM (qsu)", y= "Nitrate (mgL)", title= "Nitrate vs FDOM for All Locations")+
  geom_smooth(aes(x=FDOMqsu, y=NO3_mgL),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=FDOMqsu, y=Turb, color=Year))+
  labs(x="FDOM (qsu)", y= "Turbidity (NTU)", title= "Turbidity vs FDOM for All Locations")+
  geom_smooth(aes(x=FDOMqsu, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 30)

ggplot(data=AllData) +
  geom_point(mapping=aes(x=FDOMqsu, y=Discharge, color=Year))+
  labs(x="FDOM (qsu)", y= "Discharge (csu)", title= "Discharge vs FDOM for All Locations")+
  geom_smooth(aes(x=FDOMqsu, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=NO3_mgL, y=Turb, color=Year))+
  labs(x="Nitrate (mgL)", y= "Turbidity (NTU)", title= "Turbidity vs Nitrate for All Locations")+
  geom_smooth(aes(x=NO3_mgL, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))+
  ylim(NA, 25)

ggplot(data=AllData) +
  geom_point(mapping=aes(x=NO3_mgL, y=Discharge, color=Year))+
  labs(x="Nitrate (mgL)", y= "Discharge (csu)", title= "Discharge vs Nitrate for All Locations")+
  geom_smooth(aes(x=NO3_mgL, y=Discharge),method='lm', se=FALSE)+
  scale_color_manual(values= c("#E69F00", "#56B4E9", "#009E73", "#F0E442"))

ggplot(data=AllData) +
  geom_point(mapping=aes(x=Discharge, y=Turb, color=Year))+
  labs(x= "Discharge (csu)", y= "Turbidity (NTU)", title= "Discharge vs Turbidity")+
  geom_smooth(aes(x=Discharge, y=Turb),method='lm', se=FALSE)+
  scale_color_manual(values= c("#0072B2", "#56B4E9", "#009E73", "#D55E00"))+
  ylim(NA, 25) +
  annotate(geom="label", x=80000, y=22, label= "r = 0.12")

#"#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"
```

DFA for 4 trendlines. Diagonal and unequal used, no covariates used. Discharge as a covariate tested for 2018 SI
```{r DFA}
#Running DFA
cntl.list = list(maxit=5000, allow.degen=FALSE )
      dfa.model = list(A="zero", R="diagonal and unequal", m=4)
      four.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
  
      four.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      four.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      four.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      four.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      four.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      four.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      four.dfa.SI2018dis=MARSS(SI.2018c, model=dfa.model, control=cntl.list, covariates= Discharge8cs, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      #SAVE 'EM!
saveRDS(four.dfa.MC2015, file="MC2015DFAfour.RData")
saveRDS(four.dfa.MC2016, file="MC2016DFAfour.RData")
saveRDS(four.dfa.MC2017, file="MC2017DFAfour.RData")
saveRDS(four.dfa.MC2018, file="MC2018DFAfour.RData")
saveRDS(four.dfa.SI2015, file="SI2015DFAfour.RData")
saveRDS(four.dfa.SI2016, file="SI2017DFAfour.RData")
saveRDS(four.dfa.SI2018, file="SI2018DFAfour.RData")

#readRDS
#qqy<-readRDS( file="SI2015DFAobj.RData")
 
```    
SAMPLE to visualize things, test things, etc

```{r Sample DFA Trendline Graph}
SampTrends<- as.data.frame(t(MC.2017))
SampTrends<-SampTrends %>% mutate(T1= mean('Turb', 'FDOMqsu', 'NO3_mgL')) %>% mutate(T2=mean('CHLugL', 'BGAugL'))
SampTrends$T1= (SampTrends$Turb+SampTrends$FDOMqsu+SampTrends$NO3_mgL)/3
SampTrends$T2= (SampTrends$CHLugL+ SampTrends$BGAugL)/2
SampTrends<- SampTrends %>% mutate(num= c(58:177))

ggplot(data=SampTrends) + geom_line(mapping=aes(x=num, y=T1), color= "black", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=-1.5, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend 1 ")+
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  
  theme(axis.title.y = element_text(size = 30))   +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Trend 1")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

ggplot(data=SampTrends) + geom_line(mapping=aes(x=num, y=T2), color= "black", size=1) + 
  #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
  geom_vline(xintercept=c( 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c( 50, 80, 111, 142)+15), y=-1.5, label=c(  "June", "July", "Aug", "Sept" ), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend 2 ")+
  theme(axis.title.y = element_text(size = 30))   +
  #theme(axis.title.x = element_blank()) +
  #scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
  theme(legend.position= "none") +
  scale_y_continuous(labels=scaleFUN) +
  ggtitle("Trend 2")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
```

DFA for 3 (no covariates), 5 (no covariates), 5 (discharge as covariate) trendlines, respectively. Diagonal and unequal used. Compare the AICc values

```{r DFA for 3 and 5 trends}
dfa.model = list(A="zero", R="diagonal and unequal", m=3)
three.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
  
      three.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      three.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      three.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      three.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      three.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      three.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
      
  dfa.model = list(A="zero", R="diagonal and unequal", m=5)    
      five.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
  
      five.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      five.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
      

            five.dis.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, covariates=Discharge5c, form="dfa", z.score=TRUE, fit=TRUE)
  
      five.dis.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, covariates=Discharge6,form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      five.dis.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, covariates=Discharge7, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dis.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, covariates=Discharge8, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dis.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, covariates=Discharge5cs, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dis.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, covariates=Discharge6s, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      five.dis.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, covariates=Discharge8cs, form="dfa",  z.score=TRUE, fit=TRUE)
      #Compare AICc values
      
      #MC 2015
      print("\n 2015")
      one.dfa.MC2015$AICc
      two.dfa.MC2015$AICc
      three.dfa.MC2015$AICc
      four.dfa.MC2015$AICc
      five.dfa.MC2015$AICc #
      five.dis.dfa.MC2015$AICc
      
      print("\n 2016")
      #MC 2016
      one.dfa.MC2016$AICc
      two.dfa.MC2016$AICc
      three.dfa.MC2016$AICc
      four.dfa.MC2016$AICc
      five.dfa.MC2016$AICc #
      five.dis.dfa.MC2016$AICc
      
      #MC 2017
      print("\n 2017")
      one.dfa.MC2017$AICc
      two.dfa.MC2017$AICc
      three.dfa.MC2017$AICc
      four.dfa.MC2017$AICc
      five.dfa.MC2017$AICc #
      five.dis.dfa.MC2017$AICc
      
      #MC 2018
      print("\n 2018")
      one.dfa.MC2018$AICc
      two.dfa.MC2018$AICc
      three.dfa.MC2018$AICc
      four.dfa.MC2018$AICc
      five.dfa.MC2018$AICc #
      five.dis.dfa.MC2018$AICc
      
      #SI 2015
      print("\n 2015")
      one.dfa.SI2015$AICc
      two.dfa.SI2015$AICc
      three.dfa.SI2015$AICc
      four.dfa.SI2015$AICc
      five.dfa.SI2015$AICc #
      five.dis.dfa.SI2015$AICc
      
      #SI 2016
      print("\n 2016")
      one.dfa.SI2016$AICc
      two.dfa.SI2016$AICc
      three.dfa.SI2016$AICc
      four.dfa.SI2016$AICc #
      five.dfa.SI2016$AICc
      five.dis.dfa.SI2016$AICc
      
      #SI 2018
      print("\n 2018")
      one.dfa.SI2018$AICc
      two.dfa.SI2018$AICc
      three.dfa.SI2018$AICc
      four.dfa.SI2018$AICc
      five.dfa.SI2018$AICc #
      five.dis.dfa.SI2018$AICc
```
Run with 1 and 2 trendlines to compare
```{r 1 and 2 trends}
 dfa.model = list(A="zero", R="diagonal and unequal", m=1) 
one.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
  
      one.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      one.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      one.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      one.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      one.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      one.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
      
  dfa.model = list(A="zero", R="diagonal and unequal", m=2)    
      two.dfa.MC2015=MARSS(MC.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
  
      two.dfa.MC2016=MARSS(MC.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) #132
     
      two.dfa.MC2017=MARSS(MC.2017, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      two.dfa.MC2018=MARSS(MC.2018, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      two.dfa.SI2015=MARSS(SI.2015c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      two.dfa.SI2016=MARSS(SI.2016, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE) 
      
      two.dfa.SI2018=MARSS(SI.2018c, model=dfa.model, control=cntl.list, form="dfa",  z.score=TRUE, fit=TRUE)
      
      
 
      
      
```

Function to get the fits for each variable based on the z matrix

```{r Get fits}
get_DFA_fits <- function(MLEobj,dd=NULL,alpha=0.05) {
  fits <- list() ## empty list for results
  Ey <- MARSS:::MARSShatyt(MLEobj) ## extra stuff for var() calcs
	ZZ <- coef(MLEobj, type="matrix")$Z ## model params
	nn <- dim(Ey$ytT)[1] ## number of obs ts
	TT <- dim(Ey$ytT)[2] ## number of time steps
	H_inv <- varimax(ZZ)$rotmat ## get the inverse of the rotation matrix
	if(!is.null(dd)) {## check for covars
	  DD <- coef(MLEobj, type="matrix")$D ## model expectation
	  fits$ex <- ZZ %*% H_inv %*% MLEobj$states + DD %*% dd
	} else { ## model expectation
	  fits$ex <- ZZ %*% H_inv %*% MLEobj$states
	}
	VtT <- MARSSkfss(MLEobj)$VtT ## Var in model fits
	VV <- NULL
	for(tt in 1:TT) {
	  RZVZ <- coef(MLEobj, type="matrix")$R - ZZ%*%VtT[,,tt]%*%t(ZZ)
	  SS <- Ey$yxtT[,,tt] - Ey$ytT[,tt,drop=FALSE] %*% t(MLEobj$states[,tt,drop=FALSE])
	  VV <- cbind(VV,diag(RZVZ + SS%*%t(ZZ) + ZZ%*%t(SS)))
	  }
 	SE <- sqrt(VV)
 	fits$up <- qnorm(1-alpha/2)*SE + fits$ex ## upper & lower (1-alpha)% CI
 	fits$lo <- qnorm(alpha/2)*SE + fits$ex
 	return(fits)
}
```

Find the inverse of the Z matrix using varimax, find the rotated trendlines. Combine into Zrot4, where each line is labelled with their variable type and data set

```{r Inverse}
#MC 2015
Z.est45 = coef( four.dfa.MC2015, type="matrix")$Z
H.inv45 = varimax(Z.est45)$rotmat 
Z.rot45 = Z.est45 %*% H.inv45
trends.rot45 = solve(H.inv45) %*%  four.dfa.MC2015$states

Z.est46 = coef( four.dfa.MC2016, type="matrix")$Z
H.inv46 = varimax(Z.est46)$rotmat 
Z.rot46 = Z.est46 %*% H.inv46
trends.rot46 = solve(H.inv46) %*%  four.dfa.MC2016$states

Z.est47 = coef( four.dfa.MC2017, type="matrix")$Z
H.inv47 = varimax(Z.est47)$rotmat 
Z.rot47 = Z.est47 %*% H.inv47
trends.rot47 = solve(H.inv47) %*%  four.dfa.MC2017$states

Z.est48 = coef( four.dfa.MC2018, type="matrix")$Z
H.inv48 = varimax(Z.est48)$rotmat 
Z.rot48 = Z.est48 %*% H.inv48
trends.rot48 = solve(H.inv48) %*%  four.dfa.MC2018$states

Z.ests45 = coef( four.dfa.SI2015, type="matrix")$Z
H.invs45 = varimax(Z.ests45)$rotmat 
Z.rots45 = Z.ests45 %*% H.invs45
trends.rots45 = solve(H.invs45) %*%  four.dfa.SI2015$states

Z.ests46 = coef( four.dfa.SI2016, type="matrix")$Z
H.invs46 = varimax(Z.ests46)$rotmat 
Z.rots46 = Z.ests46 %*% H.invs46
trends.rots46 = solve(H.invs46) %*%  four.dfa.SI2016$states

Z.ests48 = coef( four.dfa.SI2018, type="matrix")$Z
H.invs48 = varimax(Z.ests48)$rotmat 
Z.rots48 = Z.ests48 %*% H.invs48
trends.rots48 = solve(H.invs48) %*%  four.dfa.SI2018$states

#Save the Z matrixes, combine them: For four trends
Z.rot45x<-as.data.frame(Z.rot45) %>% mutate(Type= spp, Year= c("MC2015"))
Z.rot46x<-as.data.frame(Z.rot46) %>% mutate(Type= spp, Year= c("MC2016"))
Z.rot47x<-as.data.frame(Z.rot47) %>% mutate(Type= spp, Year= c("MC2017"))
Z.rot48x<-as.data.frame(Z.rot48) %>% mutate(Type= spp, Year= c("MC2018"))
Z.rot45sx<-as.data.frame(Z.rots45) %>% mutate(Type= spp, Year= c("SI2015"))
Z.rot46sx<-as.data.frame(Z.rots46) %>% mutate(Type= spp, Year= c("SI2016"))
Z.rot48sx<-as.data.frame(Z.rots48) %>% mutate(Type= spp, Year= c("SI2018"))
Z.rot4<- rbind(Z.rot45x, Z.rot46x, Z.rot47x, Z.rot48x, Z.rot45sx, Z.rot46sx, Z.rot48sx)
write.csv(Z.rot4, file="Z.rot4.csv")
```

Plot the loadings for each trendline

```{r Loadings}
#Trend 1
ggplot(data= Z.rot4 ) +
  geom_linerange(mapping=aes(x=Type, ymin=0, ymax= V1, color=Year), stat= "identity", size= 1.25, position = position_dodge(width=1))+
  scale_x_discrete(labels=c("BGAugL" = "Blue-Green Algae", "CHLugL" = "Chlorophyll",
                              "FDOMqsu" = "Dissolved \n Organic Matter", "NO3_mgL"= "Nitrate", "Turb"= "Turbidity"))+
  labs(x= "Water Quality Constituent")+
  theme(panel.grid.major.x = element_blank()) +
  geom_hline(yintercept=0,color= "black") +
  geom_vline(xintercept=c(1.5, 2.5, 3.5, 4.5), color= "white", size=1.5)+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  ggtitle("Loadings for Trend One")

#Trend 2
ggplot(data= Z.rot4 ) +
  geom_linerange(mapping=aes(x=Type, ymin=0, ymax= V2, color=Year), stat= "identity", size= 1.25, position = position_dodge(width=1))+
  scale_x_discrete(labels=c("BGAugL" = "Blue-Green Algae", "CHLugL" = "Chlorophyll",
                              "FDOMqsu" = "Dissolved \n Organic Matter", "NO3_mgL"= "Nitrate", "Turb"= "Turbidity"))+
  labs(x= "Water Quality Constituent")+
  theme(panel.grid.major.x = element_blank()) +
  geom_hline(yintercept=0,color= "black") +
  geom_vline(xintercept=c(1.5, 2.5, 3.5, 4.5), color= "white", size=1.5)+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  ggtitle("Loadings for Trend Two")

#Trend 3
ggplot(data= Z.rot4 ) +
  geom_linerange(mapping=aes(x=Type, ymin=0, ymax= V3, color=Year), stat= "identity", size= 1.25, position = position_dodge(width=1))+
  scale_x_discrete(labels=c("BGAugL" = "Blue-Green Algae", "CHLugL" = "Chlorophyll",
                              "FDOMqsu" = "Dissolved \n Organic Matter", "NO3_mgL"= "Nitrate", "Turb"= "Turbidity"))+
  labs(x= "Water Quality Constituent")+
  theme(panel.grid.major.x = element_blank()) +
  geom_hline(yintercept=0,color= "black") +
  geom_vline(xintercept=c(1.5, 2.5, 3.5, 4.5), color= "white", size=1.5)+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  #facet_wrap(~Year)+
  ggtitle("Loadings for Trend Three")

#Trend Four
ggplot(data= Z.rot4 ) +
  geom_linerange(mapping=aes(x=Type, ymin=0, ymax= V4, color=Year), stat= "identity", size= 1.25, position = position_dodge(width=1))+
  scale_x_discrete(labels=c("BGAugL" = "Blue-Green Algae", "CHLugL" = "Chlorophyll",
                              "FDOMqsu" = "Dissolved \n Organic Matter", "NO3_mgL"= "Nitrate", "Turb"= "Turbidity"))+
  labs(x= "Water Quality Constituent")+
  theme(panel.grid.major.x = element_blank()) +
  geom_hline(yintercept=0,color= "black") +
  geom_vline(xintercept=c(1.5, 2.5, 3.5, 4.5), color= "white", size=1.5)+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  #facet_wrap(~Year)+
  ggtitle("Loadings for Trend Four")
```

Below, the trendlines are combined for each of the 4 trendlines. The num is the day, where the first day is the first day of collection in 2016. The trendlines are plotted (for the 4 trendlines). Each year is given its own 'type' in a new variable


```{r Trends}
ylbl<- rownames(MC.2015)
w_ts <- seq(dim(MC.2015)[2])
N_ts <- dim(MC.2015)[1]
clr <- c("brown","blue","darkgreen","darkred","purple")


#Four trends
trends.rot4<- rbind(as.data.frame(t(trends.rot45)) %>% mutate( num= c(85:171)) %>% mutate(Type= c("MC2015")), 
                   as.data.frame(t(trends.rot46)) %>% mutate( num= c(1:204)) %>% mutate(Type= c("MC2016")),
                   as.data.frame(t(trends.rot47)) %>% mutate( num= c(58:177)) %>% mutate(Type= c("MC2017")),
                   as.data.frame(t(trends.rot48)) %>% mutate( num= c(50:171)) %>% mutate(Type= c("MC2018")))
trends.rots4<- rbind(as.data.frame(t(trends.rots45)) %>% mutate( num= c(50:186)) %>% mutate(Type= c("SI2015")), 
                   as.data.frame(t(trends.rots46)) %>% mutate( num= c(1:185)) %>% mutate(Type= c("SI2016")),
                   as.data.frame(t(trends.rots48)) %>% mutate( num= c(40:178)) %>% mutate(Type= c("SI2018")))
#write.csv(trends.rot4, file="MCtrends.rot4.csv")
#write.csv(trends.rots4, file="SItrends.rot4.csv")

#Main Channel Trend 1
ggplot(data= trends.rot4) +
  geom_line(mapping= aes(x=num, y= V1, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-6, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018")) +
  ggtitle("Trend One in Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Main Channel Trend 2
ggplot(data= trends.rot4) +
  geom_line(mapping= aes(x=num, y= V2, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-6, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018")) +
  ggtitle("Trend Two in Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Main Channel Trend 3
ggplot(data= trends.rot4) +
  geom_line(mapping= aes(x=num, y= V3, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-11, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018")) +
  ggtitle("Trend Three in Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Main Channel Trend 4
ggplot(data= trends.rot4) +
  geom_line(mapping= aes(x=num, y= V4, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-11, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018")) +
  ggtitle("Trend Three in Main Channel")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Backwater Trend 1
ggplot(data= trends.rots4) +
  geom_line(mapping= aes(x=num, y= V1, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-6, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Backwater 2015", "Backwater 2016",  "Backwater 2018")) +
  ggtitle("Trend One in Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Backwater Trend 2
ggplot(data= trends.rots4) +
  geom_line(mapping= aes(x=num, y= V2, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-6, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Backwater 2015", "Backwater 2016",  "Backwater 2018")) +
  ggtitle("Trend Two in Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Backwater trend 3
ggplot(data= trends.rots4) +
  geom_line(mapping= aes(x=num, y= V3, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-7, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  ggtitle("Trend Three in Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

#Backwater trend 4
ggplot(data= trends.rots4) +
  geom_line(mapping= aes(x=num, y= V4, color= Type), size=1) +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-7, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  labs(x= "Month", y= "Trend")+
  scale_color_manual(values= c("#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  ggtitle("Trend Three in Backwater")+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
?scale_x
```

Below, the trendlines are graphed with discharge, to see the comparison between discharge and the trendlines. The correlations are found. There is a new graph for each year/location

```{r Trend and Discharge}
#Graphing discharge against each trend line, does show similarity
ggplot((data=trends.rot4 %>% filter(Type=='MC2015')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  geom_line(data=(MC.2015.rawc %>% mutate(num=85:171)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  labs(x="Time", y= "Trendline", title= "Main Channel 2015")+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rot45[1,], MC.2015.rawc$Discharge)
cor(trends.rot45[2,], MC.2015.rawc$Discharge)
cor(trends.rot45[3,], MC.2015.rawc$Discharge)
cor(trends.rot45[4,], MC.2015.rawc$Discharge)


ggplot((data=trends.rot4 %>% filter(Type=='MC2016')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Main Channel 2016")+
  geom_line(data=(MC.2016.raw %>% mutate(num=1:204)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rot46[1,], MC.2016.raw$Discharge)
cor(trends.rot46[2,], MC.2016.raw$Discharge)
cor(trends.rot46[3,], MC.2016.raw$Discharge)
cor(trends.rot46[4,], MC.2016.raw$Discharge)

ggplot((data=trends.rot4 %>% filter(Type=='MC2017')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Main Channel 2017")+
  geom_line(data=(MC.2017.raw %>% mutate(num=58:177)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")


cor(trends.rot47[1,], MC.2017.raw$Discharge)
cor(trends.rot47[2,], MC.2017.raw$Discharge)
cor(trends.rot47[3,], MC.2017.raw$Discharge)
cor(trends.rot47[4,], MC.2017.raw$Discharge)

ggplot((data=trends.rot4 %>% filter(Type=='MC2018')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Main Channel 2018")+
  geom_line(data=(MC.2018.raw %>% mutate(num=50:171)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rot48[1,], MC.2018.raw$Discharge)
cor(trends.rot48[2,], MC.2018.raw$Discharge)
cor(trends.rot48[3,], MC.2018.raw$Discharge)
cor(trends.rot48[4,], MC.2018.raw$Discharge)

ggplot((data=trends.rots4 %>% filter(Type=='SI2015')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Backwater 2015")+
  geom_line(data=(SI.2015.rawc %>% mutate(num=50:186)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rots45[1,], SI.2015.rawc$Discharge)
cor(trends.rots45[2,], SI.2015.rawc$Discharge)
cor(trends.rots45[3,], SI.2015.rawc$Discharge)
cor(trends.rots45[4,], SI.2015.rawc$Discharge)

ggplot((data=trends.rots4 %>% filter(Type=='SI2016')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Backwater 2016")+
  geom_line(data=(SI.2016.raw %>% mutate(num=1:185)), mapping=aes(x=num, y=(Discharge/5000)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rots46[1,], SI.2016.raw$Discharge)
cor(trends.rots46[2,], SI.2016.raw$Discharge)
cor(trends.rots46[3,], SI.2016.raw$Discharge)
cor(trends.rots46[4,], SI.2016.raw$Discharge)

ggplot((data=trends.rots4 %>% filter(Type=='SI2018')))+ 
  geom_line(mapping=aes(x=num, y=V1, color="#D55E00"))+
  geom_line(mapping=aes(x=num, y=V2, color="#0072B2"))+
  geom_line(mapping=aes(x=num, y=V3, color="#F0E442"))+
  geom_line(mapping=aes(x=num, y=V4, color="#009E73"))+
  labs(x="Time", y= "Trendline", title= "Backwater 2018")+
  geom_line(data=(SI.2018.rawc %>% mutate(num=40:178)), mapping=aes(x=num, y=(Discharge/7500)-4, color="black"), size=2.5)+
  guides(color = guide_legend(reverse=F,title="Line"))+
  scale_colour_identity(breaks=c("#D55E00", "#0072B2", "#F0E442", "#009E73", "black"), labels= c("Trend 1", "Trend 2", "Trend 3", "Trend 4", "Discharge"), guide="legend")

cor(trends.rots48[1,], SI.2018.rawc$Discharge)
cor(trends.rots48[2,], SI.2018.rawc$Discharge)
cor(trends.rots48[3,], SI.2018.rawc$Discharge)
cor(trends.rots48[4,], SI.2018.rawc$Discharge)

#"#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"

```
Below is a graph of the sum of the loadings for each consitutent (this is for 4 trendlines)
Addtionally, there is a graph to visualize the loadings for MC 2015 for the 3 trendlines (better graph in Chunk 7)
Loop through to find the difference between the loadings for each trendline, as well as the association (positive:both in the same direction, negative: in different directions). There are two graphs to visualize each of these. There is a third graph to combine the visualization of both
```{r Trend Loadings By Year, Asc and cor}
#Visualize the loadings for trend 1 for each location/yr for each variable
Z.rot4<-Z.rot4 %>% mutate(Sum1= (V1+V2+V3+V4))
Z.rot4<-Z.rot4 %>% select(!Sum) %>% rename(SUM=Sum1)

#SUM of the loadings for each yr/location
ggplot(data= Z.rot4 ) +
  geom_linerange(mapping=aes(x=Type, ymin=0, ymax= SUM, color=Year),  size= 1.25, position = position_dodge(width=1))+
  scale_x_discrete(labels=c("BGAugL" = "Blue-Green Algae", "CHLugL" = "Chlorophyll",
                              "FDOMqsu" = "Dissolved \n Organic Matter", "NO3_mgL"= "Nitrate", "Turb"= "Turbidity"))+
  labs(x= "Water Quality Constituent", y="Sum of Loadings for all Trendlines")+
  theme(panel.grid.major.x = element_blank()) +
  geom_hline(yintercept=0,color= "black") +
  geom_vline(xintercept=c(1.5, 2.5, 3.5, 4.5), color= "white", size=1.5)+
  scale_color_manual(values= c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"), name= "Data", labels=c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")) +
  
  
#Loops t find the differences and associations, stored in loadings.dif4
Variables= c("Turb", "CHL-a", "BGA", "FDOM", "NO3")
Trends= c(1, 2, 3, 4)
Loadings.dif4 = data.frame() #list to save model objects 
#MC2015
for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot45x[i,k]<0 & Z.rot45x[j,k]>0) | (Z.rot45x[i,k]>0 & Z.rot45x[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot45x[i,k]-Z.rot45x[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "MC2015",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop
#MC2016
for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot46x[i,k]<0 & Z.rot46x[j,k]>0) | (Z.rot46x[i,k]>0 & Z.rot46x[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot46x[i,k]-Z.rot46x[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "MC2016",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop

#MC2017
for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot47x[i,k]<0 & Z.rot47x[j,k]>0) | (Z.rot47x[i,k]>0 & Z.rot47x[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot47x[i,k]-Z.rot47x[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "MC2017",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop

#MC 2018
for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot48x[i,k]<0 & Z.rot48x[j,k]>0) | (Z.rot48x[i,k]>0 & Z.rot48x[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot48x[i,k]-Z.rot48x[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "MC2018",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop


#SI 2015
for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot45sx[i,k]<0 & Z.rot45sx[j,k]>0) | (Z.rot45sx[i,k]>0 & Z.rot45sx[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot45sx[i,k]-Z.rot45sx[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "SI2015",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop

for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot46sx[i,k]<0 & Z.rot46sx[j,k]>0) | (Z.rot46sx[i,k]>0 & Z.rot46sx[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot46sx[i,k]-Z.rot46sx[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "SI2016",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop

for(k in 1:4){ #loop over 1-4 trends
  for(i in 1:4) { #loop over the first item in the pair
    for(j in (i+1):5) {
      if((Z.rot48sx[i,k]<0 & Z.rot48sx[j,k]>0) | (Z.rot48sx[i,k]>0 & Z.rot48sx[j,k]<0)){
        asc= "Neg"
      }
      else{
        asc="Pos"
      }
      Loadings.dif4 = rbind(Loadings.dif4, 
                                     data.frame(x=abs(Z.rot48sx[i,k]-Z.rot48sx[j,k]), 
                                                Type= paste(Variables[i],Variables[j], sep="-"), 
                                                Trend= paste("V", k, sep=""),
                                                Year= "SI2018",
                                                Asc= asc))
      } #end i loop
    } # end i loop 
  }#end k loop


ggplot(data= Loadings.dif4) +
  geom_boxplot(mapping=aes(x=reorder(Type, x, median), y=x))+
  scale_x_discrete(labels=c("BGA \n FDOM", "BGA \n NO3", "CHL-a \n BGA", "CHL-a \n FDOM", "CHL-a \n NO3", "FDOM \n NO3", "Turb \n BGA", "Turb \n CHL", "Turb \n FDOM", "Turb \n NO3")) + labs(x= "Pairings", y="Absolute Value of the Difference Between Loadings", title= "Average Difference Between Loadings for Water Quality Variable Pairs")
  
  
ggplot(data= Loadings.dif4)+
  geom_bar(mapping=aes(x=reorder(Type, x, median), fill= Asc), position="stack") +
  scale_x_discrete(labels=c("BGA \n FDOM", "BGA \n NO3", "CHL-a \n BGA", "CHL-a \n FDOM", "CHL-a \n NO3", "FDOM \n NO3", "Turb \n BGA", "Turb \n CHL", "Turb \n FDOM", "Turb \n NO3")) +
   scale_fill_manual(values= c("#D55E00", "#009E73"), name= "Correlation", labels= c("Negative", "Postitive")) +
  labs(x="Pairs of Constituents", y="Count")+
  #facet_wrap(~Trend)
  ggtitle("Loadings Associations")


#Savingand combining all of the pairs info to graph
pairs<- Loadings.dif4[1:10,2] #Saving pair names
Correlations= data.frame()

Loadings.dif4[Loadings.dif4$Type=='Turb-CHL-a',]
Loadings.dif4[Loadings.dif4$Type=='Turb-CHL-a',][,5]=='Pos'
sum(Loadings.dif4[Loadings.dif4$Type=='Turb-CHL-a',][,5]=='Pos')

for(i in 1:10){
  Correlations[i,1] = pairs[i]
  Correlations[i, 2] = median(Loadings.dif4[Loadings.dif4$Type==pairs[i],][,1])
  Correlations[i, 3] =sum(Loadings.dif4[Loadings.dif4$Type==pairs[i],][,5]=='Pos')
}

colnames(Correlations)<-c("Pair", "Median", "PosCount")


#Looking at correlations of loadings
ggplot(data=Correlations) +
  geom_point(mapping=aes(x=Median, y=PosCount), color= "#0072B2", size=2)+
  geom_text(aes(Median, PosCount, label=c("Turb \n CHL-a", "Turb \n BGA", "FDOM \n Turb", "NO3 \n Turb", "CHL-a \n BGA", "FDOM \n CHL-a", "NO3 \n CHL-a", "BGA \n FDOM", "NO3 \n BGA", "NO3 \n FDOM")), nudge_x= c(.015,.008, -.015, .015, .015, .015, .015, -.015, -.015, .015), size=3.5)+ 
  labs(title= "Correlations between Loadings for Pairs of Constituents", x= "Median Difference between Loadings", y= "Count of Loadings in the Same Direction for each Pair \n (Out of 28)")+
  annotate(geom= "label", x=c(.08, .4, .35), y=c(23.25,12.5, 20), label=c("Strong \n Correlation", "Weak \n Correlation", "FDOM: Fluorescent Dissolved \n Organic Matter \n BGA: Blue-Green Algae \n CHL-a: Chlorophyll-a \n NO3: Nitrate \n Turb: Turbidity"), color= c("#D55E00", "#D55E00", "black"))

write.csv(Loadings.dif4, "Loadingsdif4.csv", row.names = FALSE)
#"#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"
```


Find the r squared value for the 4 trendlines. rqs(s)(#) is a matrix of the r squared for each variable and trendline, ie how well does each trendline predict each individual variable. rqs(s)(#)f is a matrix of the r squared for each variable and fit, ie how well does each fit found from the combination of the trendlines predict each individual variable. The s is added for backwater and the # is the year (5-8). Save these and plot them

```{r R^2 4 trends}
#MC 2015
mod5= four.dfa.MC2015
mod_fit5 <- get_DFA_fits(mod5, alpha= .05)
trends.rot5t<- t(trends.rot5)

#Transposing not necessary
#R251t = rsquare(trends.rot5t[,1], data = MC.2015r[,1])
#df51t<- data.frame(trends.rot5t[,1], MC.2015[,1])

#R sq function
rsq <- function (x, y) cor(x, y) ^ 2

mod5= four.dfa.MC2015
mod_fit5 <- get_DFA_fits(mod5, alpha= .05)
mod6= four.dfa.MC2016
mod_fit6 <- get_DFA_fits(mod6, alpha= .05)
mod7= four.dfa.MC2017
mod_fit7 <- get_DFA_fits(mod7, alpha= .05)
mod8= four.dfa.MC2018
mod_fit8 <- get_DFA_fits(mod8, alpha= .05)

mods5= four.dfa.SI2015
mod_fits5 <- get_DFA_fits(mods5, alpha= .05)
mods6= four.dfa.SI2016
mod_fits6 <- get_DFA_fits(mods6, alpha= .05)
mods8= four.dfa.SI2018
mod_fits8 <- get_DFA_fits(mods8, alpha= .05)

#rqs(s)# : r^2 for each variable and trend line
#rsq#(s)f: r^2 for each fit and variable

#MC 2015
rsq5<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq5[i,j]<- rsq(trends.rot45[j,],MC.2015.ct[i,])
  }
}
rownames(rsq5)<- spp

rsq5f<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq5f[i,j]<- rsq(mod_fit5$ex[j,],MC.2015.ct[i,])
  }
}
rownames(rsq5f)<- spp
colnames(rsq5f)<- paste(spp, "ex", sep="")


#MC 2016
rsq6<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq6[i,j]<- rsq(trends.rot46[j,],MC.2016[i,])
  }
}
rownames(rsq6)<- spp

rsq6f<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq6f[i,j]<- rsq(mod_fit6$ex[j,],MC.2016[i,])
  }
}
rownames(rsq6f)<- spp
colnames(rsq6f)<- paste(spp, "ex", sep="")

#MC 2017
rsq7<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq7[i,j]<- rsq(trends.rot47[j,],MC.2017[i,])
  }
}
rownames(rsq7)<- spp

rsq7f<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq7f[i,j]<- rsq(mod_fit7$ex[j,],MC.2017[i,])
  }
}
rownames(rsq7f)<- spp
colnames(rsq7f)<- paste(spp, "ex", sep="")

#MC 2018
rsq8<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq8[i,j]<- rsq(trends.rot48[j,],MC.2018[i,])
  }
}
rownames(rsq8)<- spp

rsq8f<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq8f[i,j]<- rsq(mod_fit8$ex[j,],MC.2018[i,])
  }
}
rownames(rsq8f)<- spp
colnames(rsq8f)<- paste(spp, "ex", sep="")


#SI 2015
rsq5s<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq5s[i,j]<- rsq(trends.rots45[j,],SI.2015.ct[i,])
  }
}
rownames(rsq5s)<- spp

rsq5sf<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq5sf[i,j]<- rsq(mod_fits5$ex[j,],SI.2015.ct[i,])
  }
}
rownames(rsq5sf)<- spp
colnames(rsq5sf)<- paste(spp, "ex", sep="")

#SI 2016
rsq6s<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq6s[i,j]<- rsq(trends.rots46[j,],SI.2016[i,])
  }
}
rownames(rsq6s)<- spp

rsq6sf<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq6sf[i,j]<- rsq(mod_fits6$ex[j,],SI.2016[i,])
  }
}
rownames(rsq6sf)<- spp
colnames(rsq6sf)<- paste(spp, "ex", sep="")


#SI 2018
rsq8s<- data.frame(c())
for(i in 1:5){
  for (j in 1:4){
    rsq8s[i,j]<- rsq(trends.rots48[j,],SI.2018.ct[i,])
  }
}
rownames(rsq8s)<- spp

rsq8sf<- data.frame(c())
for(i in 1:5){
  for(j in 1:5){
    rsq8sf[i,j]<- rsq(mod_fits8$ex[j,],SI.2018.ct[i,])
  }
}
rownames(rsq8sf)<- spp
colnames(rsq8sf)<- paste(spp, "ex", sep="")

#rqs(s)# : r^2 for each variable and trend line
#rsq#(s)f: r^2 for each fit and variable

Rsquared<- rbind(as.data.frame(t(rsq5))  %>% mutate(Type= c("MC2015")), 
                   as.data.frame(t(rsq6))  %>% mutate(Type= c("MC2016")),
                   as.data.frame(t(rsq7))  %>% mutate(Type= c("MC2017")),
                   as.data.frame(t(rsq8))  %>% mutate(Type= c("MC2018")),
                   as.data.frame(t(rsq5s))  %>% mutate(Type= c("SI2015")), 
                   as.data.frame(t(rsq6s))  %>% mutate(Type= c("SI2016")),
                   as.data.frame(t(rsq8s))  %>% mutate(Type= c("SI2018")))
Rsquared<- Rsquared%>% mutate(Trend= c('V1'))
Rsquared[,'Trend']<- c('V1', 'V2', 'V3', 'V4')

#How well does each trend line predict variable
ggplot(data=Rsquared) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=Turb, fill=Trend), position='dodge')
ggplot(data=Rsquared) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=CHLugL, fill=Trend), position='dodge')
ggplot(data=Rsquared) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=BGAugL, fill=Trend), position='dodge')
ggplot(data=Rsquared) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=FDOMqsu, fill=Trend), position='dodge')
ggplot(data=Rsquared) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=NO3_mgL, fill=Trend), position='dodge')

Rsquaredf<- rbind(as.data.frame(t(rsq5f))  %>% mutate(Type= c("MC2015")), 
                   as.data.frame(t(rsq6f))  %>% mutate(Type= c("MC2016")),
                   as.data.frame(t(rsq7f))  %>% mutate(Type= c("MC2017")),
                   as.data.frame(t(rsq8f))  %>% mutate(Type= c("MC2018")),
                   as.data.frame(t(rsq5sf))  %>% mutate(Type= c("SI2015")), 
                   as.data.frame(t(rsq6sf))  %>% mutate(Type= c("SI2016")),
                   as.data.frame(t(rsq8sf))  %>% mutate(Type= c("SI2018")))
Rsquaredf<-Rsquaredf %>% mutate (Ex= NA)
Rsquaredf[,'Ex']<- paste(spp,"ex", sep="")

#How well does each fit predict each variable
ggplot(data=Rsquaredf) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=Turb, fill=Ex), position='dodge')
ggplot(data=Rsquaredf) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=BGAugL, fill=Ex), position='dodge')
ggplot(data=Rsquaredf) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=CHLugL, fill=Ex), position='dodge')
ggplot(data=Rsquaredf) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=FDOMqsu, fill=Ex), position='dodge')
ggplot(data=Rsquaredf) + geom_bar(stat= 'identity', mapping=aes(x=Type, y=NO3_mgL, fill=Ex), position='dodge')


#FInding the overall r^2 for all of the variables and predictions
rsq2015obs<- c(MC.2015c[1,], MC.2015c[2,], MC.2015c[3,], MC.2015c[4,], MC.2015c[5,])
rsq2015exp<- c(mod_fit5$ex[1,], mod_fit5$ex[2,], mod_fit5$ex[3,], mod_fit5$ex[4,], mod_fit5$ex[5,])
rsq2015<- rsq(rsq2015obs,rsq2015exp)
#rsq2015

rsq2016obs<- c(MC.2016[1,], MC.2016[2,], MC.2016[3,], MC.2016[4,], MC.2016[5,])
rsq2016exp<- c(mod_fit6$ex[1,], mod_fit6$ex[2,], mod_fit6$ex[3,], mod_fit6$ex[4,], mod_fit6$ex[5,])
rsq2016<- rsq(rsq2016obs,rsq2016exp)
#rsq2016

rsq2017obs<- c(MC.2017[1,], MC.2017[2,], MC.2017[3,], MC.2017[4,], MC.2017[5,])
rsq2017exp<- c(mod_fit7$ex[1,], mod_fit7$ex[2,], mod_fit7$ex[3,], mod_fit7$ex[4,], mod_fit7$ex[5,])
rsq2017<- rsq(rsq2017obs,rsq2017exp)
#rsq2017

rsq2018obs<- c(MC.2018[1,], MC.2018[2,], MC.2018[3,], MC.2018[4,], MC.2018[5,])
rsq2018exp<- c(mod_fit8$ex[1,], mod_fit8$ex[2,], mod_fit8$ex[3,], mod_fit8$ex[4,], mod_fit8$ex[5,])
rsq2018<- rsq(rsq2018obs,rsq2018exp)
#rsq2018

rsq2018Sobs<- c(SI.2018c[1,], SI.2018c[2,], SI.2018c[3,], SI.2018c[4,], SI.2018c[5,])
rsq2018Sexp<- c(mod_fits8$ex[1,], mod_fits8$ex[2,], mod_fits8$ex[3,], mod_fits8$ex[4,], mod_fits8$ex[5,])
rsq2018S<- rsq(rsq2018Sobs,rsq2018Sexp)
#rsq2018S

rsq2016Sobs<- c(SI.2016[1,], SI.2016[2,], SI.2016[3,], SI.2016[4,], SI.2016[5,])
rsq2016Sexp<- c(mod_fits6$ex[1,], mod_fits6$ex[2,], mod_fits6$ex[3,], mod_fits6$ex[4,], mod_fits6$ex[5,])
rsq2016S<- rsq(rsq2016Sobs,rsq2016Sexp)
rsq2016S

rsq2015Sobs<- c(SI.2015c[1,], SI.2015c[2,], SI.2015c[3,], SI.2015c[4,], SI.2015c[5,])
rsq2015Sexp<- c(mod_fits5$ex[1,], mod_fits5$ex[2,], mod_fits5$ex[3,], mod_fits5$ex[4,], mod_fits5$ex[5,])
rsq2015S<- rsq(rsq2015Sobs,rsq2015Sexp)
#rsq2015S
```
Going through to get the r^2 value between the fits and the variables for just those specifically turb/turb, chl/chl (how well does the chl fit predict the chl data, ignores those such as how well does the bga fit line predict the chl data, which is inlcuded in the above matrix). Combine into sqall and plot

```{r R^2 specifics }
#FInding the diagonals for rsq#(s)f
#How well does turb fit predict turb, chl fit predict chl
pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq5f[i,i])
}
rsqall2015<- data.frame('r^2'= c(pp, rsq2015), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq6f[i,i])
}
rsqall2016<- data.frame('r^2'= c(pp, rsq2016), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq7f[i,i])
}
rsqall2017<- data.frame('r^2'= c(pp, rsq2017), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq8f[i,i])
}
rsqall2018<- data.frame('r^2'= c(pp, rsq2018), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq5sf[i,i])
}
rsqall2015s<- data.frame('r^2'= c(pp, rsq2015S), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq6sf[i,i])
}
rsqall2016s<- data.frame('r^2'= c(pp, rsq2016S), 'Variable'= c(spp, 'Overall'))

pp<-c()
for(i in 1:5){
  pp<-c(pp, rsq8sf[i,i])
}
rsqall2018s<- data.frame('r^2'= c(pp, rsq2018S), 'Variable'= c(spp, 'Overall'))

#NOTE THESE ARE NOT IN ORDER: need to order them
sqall<- rbind(rsqall2015  %>% mutate(Type= c("MC2015")), 
                   rsqall2016  %>% mutate(Type= c("MC2016")),
                   rsqall2017  %>% mutate(Type= c("MC2017")),
                   rsqall2018  %>% mutate(Type= c("MC2018")),
                   rsqall2015s  %>% mutate(Type= c("SI2015")), 
                   rsqall2016s  %>% mutate(Type= c("SI2016")),
                   rsqall2018s  %>% mutate(Type= c("SI2018")))
YearsAreas<- c("Main Channel 2015", "Main Channel 2016", "Main Channel 2017", "Main Channel 2018", "Backwater 2015", "Backwater 2016", "Backwater 2018")
names(YearsAreas)<- c("MC2015", "MC2016", "MC2017", "MC2018", "SI2015", "SI2016", "SI2018")
ggplot(data=sqall) + geom_bar(stat= 'identity',mapping=aes(x=Variable, y=r.2)) + labs(title= "R squared Values for Dynamic Factor Analysis Predictions", x="Water Quality Variable", y= "r squared") +  facet_wrap(.~Type, labeller= labeller( Type=YearsAreas), ncol=4) #scale_x_discrete(labels= c("Turb", "Chl-a", "BGA", "FDOM", "N03", "Overall"))  +

#Mean of all of the r^2
mean(c(rsq2015, rsq2016, rsq2017, rsq2018, rsq2015S, rsq2016S, rsq2018S))

```

The graph of the fits vs the actaul data for 2016. A different graph for each variable
```{r Fit}

#2016 MC fit for each variable-for Z-scored
ylbl<- c("Turbidity \n (NTU)", "Cholorphyll-a \n (ugL)", "Blue-green \n algae (ugL)", "Dissolved \n Organic \n Matter (qsu)", "Nitrate \n (mgL)")

#2018 Main Channel
## get model fits & CI's
mod_fit4 <- get_DFA_fits(mod6, alpha= .05)


## plot the fits
mod_fit2016<-as.data.frame(t(mod_fit6$ex))
colnames(mod_fit2016)<- c(spp)
mod_fit2016<- mod_fit2016 %>% mutate(Time= c(1:204))
M2016df<- as.data.frame(t(MC.2016))
M2016df<- M2016df %>% mutate(Time= c(1:204))

NO3m6<-ggplot(data= mod_fit2016) + 
  geom_line(mapping= aes(x= Time, y= NO3_mgL, color='black'), size=1.25)+
  labs(x="Month", y= "Nitrate \n (Z-scored)", title= "DFA Model Prediction of Nitrate in the Main Channel, 2016") +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-3, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  geom_line(data=M2016df, mapping=aes(x=Time, y=NO3_mgL, color='#999999'),size=1)+
  geom_point(data=M2016df, mapping=aes(x=Time, y=NO3_mgL, color="#0072B2"),size=1.5, shape=16 )+
  theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())+
  guides(color = guide_legend(reverse=F, title= " "))+
  scale_colour_identity(breaks=c('black', "#0072B2"), labels= c("Model", "Data"), guide="legend")

turbm6<-ggplot(data= mod_fit2016) + 
  geom_line(mapping= aes(x= Time, y= Turb), size=1.25)+
  labs(x="Month", y= "Turbidity \n (Z-scored)", title= "DFA Model Prediction in the Main Channel, 2016") +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-4, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  geom_line(data=M2016df, mapping=aes(x=Time, y=Turb), color='#999999',size=1.5)+
  geom_point(data=M2016df, mapping=aes(x=Time, y=Turb), color="#009E73",size=1.25, shape=16 )+
  theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

CHLm6<-ggplot(data= mod_fit2016) + 
  geom_line(mapping= aes(x= Time, y= CHLugL), size=1.25)+
  labs(x="Month", y= "Chlorophyll-a \n (Z-scored)", title= "DFA Model Prediction of Chlorophyll-a in the Main Channel, 2016") +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-3, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  geom_line(data=M2016df, mapping=aes(x=Time, y=CHLugL), color='#999999',size=1)+
  geom_point(data=M2016df, mapping=aes(x=Time, y=CHLugL), color="#0072B2",size=1.5, shape=16 )+
  theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

BGAm6<-ggplot(data= mod_fit2016) + 
  geom_line(mapping= aes(x= Time, y= BGAugL), size=1.25)+
  labs(x="Month", y= "Blue-Green Algae \n(Z-scored)", title= "DFA Model Prediction of Blue-Green Algae in the Main Channel, 2016") +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-2.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  geom_line(data=M2016df, mapping=aes(x=Time, y=BGAugL), color='#999999',size=1)+
  geom_point(data=M2016df, mapping=aes(x=Time, y=BGAugL), color="#0072B2",size=1.5, shape=16 )+
  theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())

FDOMm6<-ggplot(data= mod_fit2016) + 
  geom_line(mapping= aes(x= Time, y= FDOMqsu), size=1.25)+
  labs(x="Month", y= "FDOM \n (Z-scored)", title= "DFA Model Prediction of Fluorescent Dissolved Organic Matter") +
  geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
  annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171)+15), y=-3.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct"), size=3.5) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank())+
  geom_line(data=M2016df, mapping=aes(x=Time, y=FDOMqsu), color='#999999',size=1)+
  geom_point(data=M2016df, mapping=aes(x=Time, y=FDOMqsu), color="#0072B2",size=1.5, shape=16 )+
   theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
  
 turbm6
 CHLm6
 BGAm6
 FDOMm6
 NO3m6

# for(i in 1:5) {
#  # up <- mod_fit$up[i,]
#   mn <- mod_fit$ex[i,]
#   #lo <- mod_fit$lo[i,]
#   plot(w_ts,mn,xlab="Time",ylab=ylbl[i],xaxt="n",type="n", cex.lab=1.2,
#        ylim=c(min(lo),(max(up))))
#   points(w_ts,MC.2018[i,], pch=16, col=clr[i])
#   #lines(w_ts, up, col="darkgray")
#   lines(w_ts, mn, col="black", lwd=2)
#   #lines(w_ts, lo, col="darkgray")
# }
# 
# AllDataMC6<-AllData %>% filter(Type=='MC2016')
# ggplot(data=AllDataMC6) + geom_line(mapping=aes(x=num, y=CHLugL), color= "#56B4E9", size=1) + 
#    geom_line(mapping=aes(x=num, y=BGAugL*10), color= "#E69F00", size=1) +
#   #facet_wrap(~Area)+ #, labeller= labeller( Type=YearsAreas)) +
#   geom_vline(xintercept=c(19, 50, 80, 111, 142, 171, 203) , col="darkgray")+
#   annotate(geom= "text", x= (c(-13,19, 50, 80, 111, 142, 171, 203)+15), y=-.5, label=c("April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov"), size=3.5) +
#   theme(panel.grid.major.x = element_blank()) +
#   theme(panel.grid.minor.x = element_blank())+
#   #labs(x= "Month", y= "Nitrate (mgL)")+
#  # scale_color_manual(values= c("#E69F00", "#56B4E9"), name= "Year", labels=c("2015", "2016", "2017", "2018")) +
#   scale_y_continuous(
# 
#     name = "Chlorophyll-a (ugL)",
#     
#     # Add a second axis and specify its features
#     sec.axis = sec_axis(~./10, name="Blue-Green Algae (ugL)")
#   ) + 
#   
#   ggtitle("Chlorophyll-a and Blue-Green Algae Levels in 2016 in the Main Channel")+
#   theme(
#     axis.title.y = element_text(color = "#56B4E9", size=13),
#     axis.title.y.right = element_text(color = "#E69F00", size=13)
#   ) +
#    theme(axis.text.x= element_blank(), axis.ticks.x = element_blank())
#"#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"
```

