---
title: 'Preliminary Exploration: Sensors'
author: "Taryn Waite"
date: "6/4/2020"
output: html_document
---

This document contains my preliminary exploration of the water quality sensor data and discharge data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dataRetrieval)
library(tidyverse)
library(lubridate)
```

## Sensor Data Wrangling

Reading in the data
```{r read data}
# main channel WQ data
mc <- read_csv("data/MC_WQ.csv")
# backwater WQ data
si <- read_csv("data/SI_WQ.csv")

# discharge data
discharge <- readNWISdv(siteNumbers = "05378500", parameterCd="00060", startDate = "2015-06-30", endDate = "2018-10-12")

```

Here, I convert the date column into Date objects and also create a date-time column
```{r date and time}
mc <- mc %>% 
  mutate(date = mdy(`YYYY-MM-DD`)) %>% 
  mutate(dateTime = mdy_hms(paste(`YYYY-MM-DD`, as.character(`hh:mm`))))

si <- si %>% 
  mutate(date = ymd(`YYYY-MM-DD`)) %>% 
  mutate(dateTime = ymd_hms(paste(`YYYY-MM-DD`, as.character(`hh:mm`))))
```


Below, I combine the SI and MC data into one dataframe with a column called "site" taking the value "SI" or "MC".
```{r combining si and mc}
# first we need a new column specifying if si or mc
si_with_ID <- si %>% mutate(site = "SI")
mc_with_ID <- mc %>% mutate(site = "MC")

# get rid of original date and time columns
# (don't need these anymore with new date and dateTime columns)
si_with_ID <- si_with_ID %>% select(-c(1,2, 23))
mc_with_ID <- mc_with_ID %>% select(-c(1,2, 23))

# next we can merge the two data frames by row
all_WQ <- bind_rows(si_with_ID, mc_with_ID)
```

## Water Quality Data Exploration 

To facilitate data exploration, below I've defined a function that plots a time series of a given variable for both sites (either on one plot or side-by-side). A year and/or month and/or day can be specified; otherwise, all time series data will be plotted.
```{r time series plot function}
# plots time series of a given variable in the WQ dataframe
# var (string): variable name to be plotted
# year, month, day (optional, numeric): year, month, and day to be plotted (can specify all, some, or none)
# if month is specified but not year, separate plots are generated for each year
# sep (boolean): plots MC and SI on same plot if false or side-by-side plots if true
plot_time_series_WQ <- function(var, year = NA, month = NA, day = NA, sep = F){
  # filter for specified year
  if(!is.na(year)){
    all_WQ <- all_WQ %>% filter(year(date) == year)
  }
  # filter for specified month
  if(!is.na(month)){
    all_WQ <- all_WQ %>% filter(month(date) == month)
  }
  # filter for specified day
  if(!is.na(day)){
    all_WQ <- all_WQ %>% filter(day(date) == day)
  }
  # if sep argument is true, plot with facet wrap
  # otherwise, plot without facet wrap
  if(sep & is.na(year) &!is.na(month)){
    all_WQ %>% 
    ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
    geom_line() +
    facet_grid(site ~ as.factor(year(date)), scales = "free_x")
  } else if (sep){
    all_WQ %>% 
    ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
    geom_line() +
    facet_wrap(~site)
  } else if (is.na(year) &!is.na(month)){
    all_WQ %>% 
    ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
    geom_line() +
    facet_wrap(~as.factor(year(date)), scales = "free_x")
  } else{
    all_WQ %>% 
    ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
    geom_line()
  }
  # if(sep){
  #   all_WQ %>% 
  #   ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
  #   geom_line() +
  #   facet_wrap(~site)
  # } else {
  #   all_WQ %>% 
  #   ggplot(aes_string(x = "dateTime", y = var, col = "site")) +
  #   geom_line()
  # }
}


```

```{r plotting function examples}
# plot of temperature for June 2016
plot_time_series_WQ("Temp", year = 2016, month = 6)

# separate plots of nitrate for all of 2016
plot_time_series_WQ("NO3_mgL", year = 2016, sep = T)

# plot of turbidity for June 9, 2016
plot_time_series_WQ("Turb", year = 2016, month = 6, day = 9)

# plots of temperature in the month of July for all years at both sites
plot_time_series_WQ("Temp", month = 7, sep = T)
```


## Merging WQ and discharge data

Below, I merge the WQ and discharge data into one dataframe that has average values of WQ variables for each day, as well as the discharge for the day.
```{r}
# group water quality data by date, giving dataframe with mean values for each day
# here, I just included variables we will likely need, but may need to add more later
daily_WQ <- all_WQ %>%
  group_by(date) %>% 
  summarise (date = mean(date), Turb = mean(Turb), CHLugL = mean(CHLugL), BGAugL = mean(BGAugL),
             FDOMqsu = mean(FDOMqsu), NO3_mgL = mean(NO3_mgL))

# select just date and discharge from discharge data, and rename date column to match WQ
date_discharge <- discharge %>% 
  select(c(3, 4)) %>% 
  rename(date = Date)

# join daily WQ and discharge data and rename discharge column
WQ_discharge <- full_join(daily_WQ, date_discharge, by = "date") %>% 
  rename(discharge = X_00060_00003)


```

Below, I've defined a function for looking at just the discharge data. It plots a time series of discharge and, like with the WQ time series plotting function, you can specify a year and/or month to plot (otherwise, all data is plotted). If you specify a month but not a year, the time series for the given month will be shown on separate plots for each year.
```{r discharge time series function}
# plots the discharge over time
# year, month (optional, numeric) specify the specific year and/or month to plot (if not specified, all data will be plotted)
# if month is given but not year, all years will be plotted separately.
plot_time_series_discharge <- function(year = NA, month = NA){
  # filter for specified year
  if(!is.na(year)){
    WQ_discharge <- WQ_discharge %>% filter(year(date) == year)
  }
  # filter for specified month
  if(!is.na(month)){
    WQ_discharge <- WQ_discharge %>% filter(month(date) == month)
  }
  
  # if month is specified but not year, plot each year separately
  # otherwise, just make one plot
  if(is.na(year) & !is.na(month)){
    WQ_discharge %>% 
      ggplot(aes(x = date, y = discharge)) +
      geom_line() +
      facet_wrap(~as.factor(year(date)), scales = "free_x")
  } else {
  WQ_discharge %>% 
    ggplot(aes(x = date, y = discharge)) +
    geom_line()
  }
}
```

Here are some examples of how the discharge time series plotting function is used:
```{r testing discharge plotting function}
# plot of all discharge data over time
plot_time_series_discharge()

# plot of discharge data from 2016
plot_time_series_discharge(year = 2016)

# plot of discharge data from June 2016
plot_time_series_discharge(year = 2016, month = 6)

# plots of discharge data from July of each year
plot_time_series_discharge(month = 7)
```

