---
title: "Exploring Phytos Data James"
author: "James Pack"
date: "6/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction

This is my exploration of the phytoplankton data. I've used the same code as my data cleaning document in the repository to set up the data, seen below.

#Data Cleaning

```{r cleaning, echo=F}
# Phytoplankton Data Cleaning James
library(tidyverse)
library(dplyr)
library(lubridate)

phytos_9606 <- read_csv('Phytos_1996-2006.csv')
phytos_0812 <- read_csv('Phytos_2008-2012.csv')
phyto_info <- read_csv('Phyto_info_WQdata.csv')

# Combining data from all time periods
phytos_whole <- phytos_9606 %>% bind_rows(phytos_0812)

# Adding water quality data, joined by SHEETBAR (ID) of sample
phyto <- phytos_whole %>% left_join(phyto_info,by="SHEETBAR")

# Cleaning the dates
phyto$DATE <- phyto$DATE %>% mdy()

# Seeing the table
phyto %>% head()
```

#Exploring the Data

To start exploration, I'm using plots to see trends.
##Trends over time
```{r tallytime, echo=FALSE, warning=FALSE}
library(ggplot2)
library(ggthemes)
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(DATE,`TOTAL BV_um3L`/1e09)) +
  geom_point(aes(color=DIVISION)) +
  facet_wrap(~FLDNUM, nrow=3) +
  ggtitle('Biomass Over Time') +
  xlab('Date') +
  ylab('Total Biomass')
```

I used different locations in the field to see if there were any noticeable trends unique to each data collection point. Aside from the outliers, it seems as though there was an increase in phytoplankton overall in the later period of the data collection, especially in stations 1, 2, and 3. The spikes were all in Bacillariophyta, so it will be interesting to see trends in other Divisions if Bacillariophyta are removed. 

```{r tallytime, echo=FALSE, warning=FALSE}
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  filter(DIVISION != 'Bacillariophyta') %>% 
  ggplot(aes(DATE,`TOTAL BV_um3L`/1e09)) +
  geom_point(aes(color=DIVISION)) +
  facet_wrap(~FLDNUM, nrow=3) +
  ggtitle('Biomass Over Time') +
  xlab('Date') +
  ylab('Total Biomass')
```

I also want to compare the trend of tally and temperature to see if there is a noticeable impact there.
##Trends with temperature
```{r tallytemp, echo=FALSE, warning=FALSE}
phyto %>% ggplot(aes(TEMP,`TOTAL BV_um3L`/1e09)) +
  geom_point(aes(color=DIVISION)) +
  geom_smooth(method=lm) +
  ggtitle('Biomass Response to Temperature') +
  xlab('Temperature (Celsius)') +
  ylab('Biomass') +
  theme_gdocs()
```

This plot doesn't emphasize just how many points are on the lower end, so while it seems that higher temperatures have a trend to decrease biomass, adding the line of best fit better shows that the trend is not noticeable. Recreating the same plot but without the Bacillariophyta that have considerably higher observations than the other divisions, you get

```{r tallytemp, echo=FALSE, warning=FALSE}
phyto %>% 
  filter(DIVISION != 'Bacillariophyta') %>% 
  ggplot(aes(TEMP,`TOTAL BV_um3L`/1e09)) +
  geom_point(aes(color=DIVISION)) +
  geom_smooth(method=lm) +
  ggtitle('Biomass Response to Temperature') +
  xlab('Temperature (Celsius)') +
  ylab('Biomass') +
  theme_gdocs()
```

Now there are almost no observable trends, with only a few outliers present. This suggests that Bacillariophyta are more affected by temperatures than the other divisions, so flipping the filter the opposite way may show a trend.

```{r tallytemp, echo=FALSE, warning=FALSE}
phyto %>% 
  filter(DIVISION == 'Bacillariophyta') %>% 
  ggplot(aes(TEMP,`TOTAL BV_um3L`/1e09)) +
  geom_point(aes(color=DIVISION)) +
  geom_smooth(method=lm) +
  ggtitle('Biomass Response to Temperature') +
  xlab('Temperature (Celsius)') +
  ylab('Biomass') +
  theme_gdocs()
```

The trend here is still not pronounced because of just how many points are near the bottom of the plot, but it is more noticeable than in the prior two plots. After removing outliers, this may be worth revisiting. 

This time, rather than differing locations, I used color to differentiate between divisions in order to see if different divisions responded differently to water temperature. Cyanobacteria seemed to greatly favor warmer water temperatures, so I'm going to remove them from the data and plot all other points to see if there's a trend.
```{r no_cyano_temp, echo=F, warning=F}
phyto %>% 
  filter(DIVISION != 'Cyanobacteria') %>% 
  ggplot(aes(TEMP,`TOTAL BV_um3L`)) +
  geom_point(aes(color=DIVISION)) +
  ggtitle('Tally Response to Temperature') +
  xlab('Temperature (Celsius)') +
  ylab('Tally') +
  theme_gdocs()
```

After removing Cyanobacteria, two trends emerged that I could see. Bacillariophyta seemed to much prefer the colder water temperatures, while Cryptophyta experienced more of an increase as water temperature increased. Chlorophyta seemed to be relatively constant with no visible variation, and Chrysophyta, Euglenophyta, and Pyrrophyta are barely observable at all.

To focus on the two divisions I saw any sort of trend in (Bacillariophyta and Cryptophyta), I am going to plot them again by temperature separately to see if anything  interesting occurs. I'm also going to add a line to the plot to see what relationship is observable.
```{r three_trends, echo=F, warning=F}
phyto %>% 
  filter(DIVISION == c('Bacillariophyta','Cryptophyta')) %>% 
  ggplot(aes(TEMP,`TOTAL BV_um3L`)) +
  geom_point() +
  facet_wrap(~DIVISION,nrow=2) +
  geom_smooth(method=lm) +
  ggtitle('Biomass Response to Temperature') +
  xlab('Temperature (Celsius)') +
  ylab('Tally') +
  theme_gdocs()
```

Somewhat surprisingly, although there was some variation in each, there was next to no observable effect of temperature on the tally observed. Bacillariophyta seems to have a slight downward trend with an increase in temperature, but not in any significant way. 

##Trends with latitude
Something that may be worth noting is that the latitude (using Northing as the variable) may have an effect on the tally. Like before, Cyanobacteria are filtered out because of how numerous they've become.
```{r latitude_trend, echo=F, warning=F}
phyto %>% 
  filter(DIVISION != 'Cyanobacteria') %>% 
  ggplot(aes(NORTHING,`TOTAL BV_um3L`)) +
  geom_point(aes(color=DIVISION)) +
  ggtitle('Tally Response to Northing') +
  xlab('Northing') +
  ylab('Tally') +
  theme_gdocs()
```

Interestingly, Bacillariophyta seem to slightly increase in tally as the northing increases. This may help explain the slight trend observed when comparing the trend caused by temperature.

The next graphic focuses solely on Bacillariophyta.
```{r bacill_trend, echo=F, warning=F}
phyto %>% 
  filter(DIVISION == 'Bacillariophyta') %>% 
  ggplot(aes(NORTHING,TALLY)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Bacillariophyta Response to Northing') +
  xlab('Northing') +
  ylab('Tally') +
  theme_gdocs()
```

This has produced the same slight trend observed before, but in the opposite direction (the more north you are the colder). Interesting, but not particularly significant. Probably a good time to move on to water quality measures instead to see if interesting trends emerge.

#Water Quality Trends
##pH Responses
```{r pH_trend, echo=F, warning=F}
phyto %>%
  filter(is.na(DIVISION) == F) %>% 
  ggplot(aes(PH,TALLY)) +
  geom_point(aes(color=DIVISION)) +
  facet_wrap(~DIVISION,nrow=3) +
  ggtitle('Tally Response to pH') +
  xlab('pH') +
  ylab('Tally') +
  theme_gdocs()
```

Filtering out Cyanobacteria because of the high tally counts, you get

```{r pH_no_cyan, echo=F, warning=F}
phyto %>%
  filter(is.na(DIVISION) == F) %>% 
  filter(DIVISION != 'Cyanobacteria') %>% 
  ggplot(aes(PH,TALLY)) +
  geom_point(aes(color=DIVISION)) +
  facet_wrap(~DIVISION,nrow=3) +
  ggtitle('Tally Response to pH') +
  xlab('pH') +
  ylab('Tally') +
  theme_gdocs()
```

This shows trends in Bacillariophyta, Chlorophyta, and Cryptophyta that were not observable with the Cyanobacteria graphic present. Now, examining these three divisions only and adding a line of best fit:
```{r pH_interesting_trends, echo=F, warning=F}
phyto %>%
  filter(is.na(DIVISION) == F) %>% 
  filter(DIVISION == c('Bacillariophyta','Chlorophyta','Cryptophyta')) %>% 
  ggplot(aes(PH,TALLY)) +
  geom_point(aes(color=DIVISION)) +
  facet_wrap(~DIVISION,nrow=3) +
  geom_smooth(method=lm) +
  ggtitle('Tally Response to pH') +
  xlab('pH') +
  ylab('Tally') +
  theme_gdocs()
```

Much like the Cyanobacteria samples, Tally increases slightly with each of these divisions as pH increases. 

#Cyanobacteria
So far in everything I've done, Cyanobacteria as a division has severely limited the amount I can learn about the others because of the frequency. Instead, I'm going to go ahead and focus on it to see if I can learn something interesting about the genera observed within the division, and how they differ.

##Over Time
```{r cyano_time, echo=F, warning=F}
phyto %>% 
  filter(DIVISION == 'Cyanobacteria') %>% 
  ggplot(aes(DATE,TALLY)) +
  geom_point() +
  facet_wrap(~GENUS,nrow=3) +
  ggtitle('Cyanobacteria over time') +
  xlab('Date') +
  ylab('Tally') +
  theme_gdocs()
```

Aside from being awful to read, the plot does show that only a few genera changed over time, while several were hardly observed at all. Now to redo the plot for a few specific genera that had over 1500 observations.
```{r cyano_filter_time, echo=F, warning=F}
phyto %>% 
  filter(DIVISION == 'Cyanobacteria') %>% 
  filter(GENUS == 'Planktolyngbya limnetica' | GENUS == 'Aphanocapsa delicatissima' | GENUS == 'Aphanizomenon sp.' | GENUS == 'Merismopedia tenuissima' | GENUS == 'Microcystis aeruginosa') %>% 
  ggplot(aes(DATE,TALLY)) +
  geom_point() +
  facet_wrap(~GENUS,nrow=2) +
  geom_smooth(method=lm) +
  ggtitle('Cyanobacteria over time') +
  xlab('Date') +
  ylab('Tally') +
  theme_gdocs()
```

Just to see, unsure how I can rewrite the above plot without having to know each of the genera names. Going to leave that as a project for tomorrow.