---
title: "Abundance vs Richness"
author: "James Pack"
date: "6/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning=F)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggthemes)
library(dplyr)
```

# Introduction

In this file, I will explore the relationship between the abundance of phytoplankton and the richness of the phytoplankton communities. While before I used Total Biovolume to measure abundance, this allows for multiple recordings of abundance at each of the six stations. While this is important and will be explored first, the end goal is to have a single measure of abundance (total biovolume across taxa) for each field station for each date of observations.

```{r data_clean, include = F}
phytos_9606 <- read_csv('Phytos_1996-2006.csv')
phytos_0812 <- read_csv('Phytos_2008-2012.csv')
phyto_info <- read_csv('Phyto_info_WQdata.csv')

# Combining data from all time periods
phytos_whole <- phytos_9606 %>% bind_rows(phytos_0812)

# Adding water quality data, joined by SHEETBAR (ID) of sample
phyto <- phytos_whole %>% left_join(phyto_info,by="SHEETBAR")

# Cleaning the dates
phyto <- phyto %>% mutate(DATE=mdy(DATE))
```

# Calculating Total Richness and Abundance

After cleaning the data using code I wrote last week, I now want to add on calculated total abundance and total richness to my data. This will allow for comparisons of these total amounts as well as for an exploration of how much biomass of a given observation certain phytoplankton make up (i.e. seeing how the relative abundance of cyanobacteria has changed over time at station 1). First, adding the richness calculation that I created earlier this week.

```{r rich, echo=FALSE}
phyto <- left_join(phyto,phyto %>% count(SHEETBAR),by='SHEETBAR') %>% 
  rename(rich=n)
```

Next is time for calculating the total abundance. Similar to the richness calculation, this value should be the same for each observation at the same location on the same date. However, this cannot be measured by counting the SHEETBAR variable. Instead, the sum of the TOTAL BV_um3L values from each location and date must be assigned to the observations. 

```{r abund}
#sum((phyto %>% filter(SHEETBAR == 41009893))$`TOTAL BV_um3L`)
total <- phyto %>% group_by(SHEETBAR) %>% 
  summarize(total_abundance = sum(`TOTAL BV_um3L`))
phyto <- left_join(phyto,total,by='SHEETBAR')
head(phyto$total_abundance)
```

# Using the Data

Now that the total abundance has been calculated, time to plot each of richness and abundance over time at all stations before comparing them.

```{r over_time}
phyto %>% ggplot(aes(DATE,total_abundance / 1e09)) +
  geom_point() + 
  geom_smooth(method=lm) +
  ggtitle('Total Abundance Over Time') +
  xlab('Date') +
  ylab('Total Biovolume') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Richness Over Time') +
  xlab('Date') +
  ylab('# of Phytoplankton Taxa Observed') +
  theme_gdocs()
```

It seems that while total abundance over time has changed, richness has only grown more varied. Plotting the two against one another may show a pattern, but comparisons over time don't give much insight on the large scale.

```{r abund_v_rich}
phyto %>% ggplot(aes(total_abundance/1e09,rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Richness vs. Abundance') +
  xlab('Total Abundance') +
  ylab('Richness') +
  theme_gdocs()
```

This yields only a slightly positive trend, which is dubious because of how few points have such a high abundance. Repeating the same code but separating based on location yields the following plots:

```{r abund_v_rich_2}
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(total_abundance/1e09,rich)) +
  geom_point() +
  facet_wrap(~FLDNUM,nrow=3) +
  geom_smooth(method=lm) +
  ggtitle('Richness vs. Abundance') +
  xlab('Total Abundance') +
  ylab('Richness') +
  theme_gdocs()
```

The same problem as before emerges, where more extreme values may be affecting the trends that are seen otherwise. 

# Blue in Green

Because increases in Cyanobacteria populations can cause harmful algal blooms, it may be important to see if the relative abundance of Cyanobacteria positively or negatively affects the richness and total abundance of the larger phytoplankton population. 

```{r cyan_abund}
rel_cyan <- phyto %>% filter(DIVISION == 'Cyanobacteria') %>% 
  group_by(SHEETBAR) %>% 
  summarize(cyan_abund = sum(`TOTAL BV_um3L`))
phyto <- left_join(phyto,rel_cyan,by='SHEETBAR')
head(phyto$cyan_abund)
```

It appears that the join has worked for all but a few observations (175 observations of cyanobacteria abundance vs. 181 observations of total abundance). Now to plot this and see how it compares to total abundance and richness of communities.

## Getting a feel for Cyanobacteria

```{r rel_cyan}
phyto %>% ggplot(aes(DATE,cyan_abund/1e9)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Cyanobacteria Abundance Over Time') +
  xlab('Date') +
  ylab('Cyanobacteria Biovolume') +
  theme_gdocs()
```

Interestingly, while there have been spikes of large masses of cyanobacteria, this does not seem to be the case more generally. In fact, despite the number of high values, the biovolume of cyanobacteria has decreased over time. 

## Cyanobacteria and the phytoplankton community

```{r cyan_tot_abund}
phyto %>% ggplot(aes(cyan_abund/1e9,total_abundance/1e9)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Total Abundance vs. Cyanobacteria Abundance') +
  xlab('Cyanobacteria Biovolume') +
  ylab('Total Biovolume') +
  theme_gdocs()
```

While there's not a definitive trend downwards as I expected, it does seem that the observations with the highest total biovolume tended to have little to no Cyanobacteria observed, while observations with high Cyanobacteria biovolume were more limited in their total biovolume. This makes sense, as there would necessarily be some sort of positive trend unless Cyanobacteria were killing other species at the same rate they reproduced or faster. Next, comparing it to richness might show some of the negative effects Cyanobacteria have.

```{r cyan_rich}
phyto %>% ggplot(aes(cyan_abund/1e9,rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Richness vs. Cyanobacteria Abundance') +
  xlab('Cyanobacteria Biovolume') +
  ylab('Richness') +
  theme_gdocs()
```

Interestingly, there is another positive trend here. It seems that Cyanobacteria have not had the clear trends I expected of them, and some more careful analysis is needed. It may interesting to separately study quartiles of the Cyanobacteria (i.e. study the observations with the highest Cyanobacteria biovolume, lowest, and middle quarters). I'm also going to consider the % of total biovolume that is Cyanobacteria compared to abundance to see if the above graphic changes.

```{r cyan_rich2}
phyto %>% ggplot(aes(cyan_abund/total_abundance*100,rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Richness vs. Cyanobacteria Abundance') +
  xlab('Cyanobacteria Biovolume %') +
  ylab('Richness') +
  theme_gdocs()
```

Despite the points seeming to have a more generally negative trend, the line drawn by ggplot was still slightly positive. By eye, it seems that is flatter than the previous plot, so this may be closer to understanding the negative effects Cyanobacteria may have. Exploring quartiles is likely the next step. 

## Location specific Cyanobacteria and phytoplankton communities

It may be worth examining more closely the growth of Cyanobacteria at each field station & the impact on local phytoplankton, as more clear patterns may emerge there.

```{r cyan_loc}
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(FLDNUM,cyan_abund/1e9)) +
  geom_point() +
  ggtitle('Cyanobacteria Abundance') +
  xlab('Field Number') +
  ylab('Cyanobacteria Biovolume') +
  theme_gdocs()
```

Maybe unsurprisingly, each location had its peak at a different point of time, if at all. Stations 1, 2, and 3 all had relatively high peaks between 6 and 8 billion, while stations 4, 5, and 6 never passed 3 billion. Focusing on stations with more change (1,2,3) may give more insights as to what allows Cyanobacteria to thrive, while stations with less change (4,5,6) may give more insight as to what inhibits or doesn't allow the growth of Cyanobacteria. It's also important to keep in mind that while stations 1, 2, and 3 did all have high points in Cyanobacteria biovolume, these observations seem to be outliers.

```{r cyan_spikes}
high_cyan <- phyto %>% filter(FLDNUM %in% c(1,2,3))
high_cyan %>% ggplot(aes(cyan_abund/1e9,total_abundance/1e9)) +
  geom_point() +
  geom_smooth(method=lm)
high_cyan %>% 
  ggplot(aes(cyan_abund/total_abundance * 100,total_abundance/1e9)) +
  geom_point() +
  geom_smooth(method=lm)
```

The first graphic didn't feel quite right, so I made the second to see if the proportion of cyanobacteria to other phytoplankton was more impactful. The second graphic seems to show closer to what I expected to see based on the reading done. As the relative abundance of cyanobacteria increases (cyanobacteria biovolume/total biolume), the total abundance tends to fall at field stations that experienced more variation. Now to see if the same holds true for richness.

```{r cyan_spikes2}
high_cyan %>% ggplot(aes(cyan_abund/1e9,rich)) +
  geom_point() +
  geom_smooth(method=lm)
high_cyan %>% 
  ggplot(aes(cyan_abund/total_abundance * 100,rich)) +
  geom_point() +
  geom_smooth(method=lm)
```

Unlike with total abundance, richness did not show a clear negative response to an increased proportion of cyanobacteria in the total biovolume measured. However, it is notable that for observations where cyanobacteria make up over 50% of the phytoplankton community, the observed measure of richness was lower than the predicted richness by the line of best fit. 

Below, I've included plots of all the field stations comparing the relative cyanobacteria abundance (cyanobacteria biovolume/total biovolume) with the richness and abundance. This will likely result in less clear trends, as stations 4, 5, and 6 experienced fewer spikes in cyanobacteria populations (as mentioned earlier).

```{r}
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(cyan_abund/total_abundance*100,total_abundance/1e9)) +
  geom_point() +
  geom_smooth(method=lm)
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(cyan_abund/total_abundance*100,rich)) +
  geom_point() +
  geom_smooth(method=lm)
```

Interestingly, only the comparison with richness showed another positive trend. It seems that while a larger proportion of cyanobacteria might reduce the overall size of phytoplankton communities, it does not in the same way reduce the richness of the community. Comparing the effect of Cyanobacteria on other taxa would be another good step in understanding the impact. 

# Cyanobacteria & Nutrients
After seeing the article Kathi Jo sent us (6/17), I wanted to test the relationship of Cyanobacteria abundance and various nutrients in the water (especially phosphorous, as mentioned in the article). 

```{r cyan_phos}
phyto %>% ggplot(aes(TP,cyan_abund/1e9)) +
  geom_point() + 
  geom_smooth(method=lm) +
  ggtitle('Total Phosphorous and Cyanobacteria Abundance') +
  xlab('Total Phosphorous') +
  ylab('Cyanobacteria Biovolume') +
  theme_gdocs()
```

Interestingly, this also does not reveal the shape I expected. However, the low Cyanobacteria biovolume observations with higher total phosphorous may be due to lower overall biovolume. This time, I will instead use the percentage of total biovolume made up by cyanobacteria against the total phosphorous available.

```{r cyan_phos2}
phyto %>% ggplot(aes(TP,cyan_abund/total_abundance * 100)) +
  geom_point() + 
  geom_smooth(method=lm) +
  ggtitle('Total Phosphorous and % Cyanobacteria Abundance') +
  xlab('Total Phosphorous') +
  ylab('Cyanobacteria Biovolume %') +
  theme_gdocs()
```

Just for thoroughness, next investigating Cyanobacteria and Nitrogen.

```{r cyan_nit}
phyto %>% ggplot(aes(TN,cyan_abund/1e9)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Total Nitrogen and Cyanobacteria Abundance') +
  xlab('Total Nitrogen') +
  ylab('Cyanobacteria Biovolume') +
  theme_gdocs()
```

And again, with relative Cyanobacteria Abundance.

```{r cyan_nit2}
phyto %>% ggplot(aes(TN,cyan_abund/total_abundance * 100)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Total Nitrogen and Cyanobacteria Relative Abundance') +
  xlab('Total Nitrogen') +
  ylab('% Cyanobacteria of Biovolume') +
  theme_gdocs()
```

Neither of these plots indicate a strong relationship between the abundance of Cyanobacteria and nutrients available. This doesn't mean necessarily that there is no relationship, but as of now I am not seeing a clear trend.