---
title: "Identifying Storm Events"
author: "Taryn Waite"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Below, I tried identifying each day as a "storm" day or a "baseline" day depending on the change in discharge between the day and the next day. A day is considered "storm" if the discharge increases by more than 10% in the day. However, I don't think this worked well, judging by the plot I created at the end. It mostly just catches the very beginning of what seem to be storm events.

```{r}

## tried with dplyr but I couldn't get it to work
dates <- discharge[[3]]

get_discharge <- function(date){
  if(date %in% dates){
    day_data <- discharge %>% 
      filter(Date == date)
    return((day_data[[4]]))
  }
  else{
    print("next day NA")
    return(NA)
  }
}


get_change_discharge <- function(date){
  return(get_discharge(date + 1 ) - get_discharge(date)) 
}

## this is where it wouldn't work 
#discharge %>%
  #mutate(change = get_change_discharge(Date))


## instead, I just found the difference by arranging by date and
## subtracting each from the previous (this works because no days are skipped)

# arrange by date
temp <- discharge %>% 
  rename(daily_discharge = X_00060_00003) %>% 
  arrange(Date) 

# calculate difference column
temp$diff <- lead(temp$daily_discharge, 1) - temp$daily_discharge

# give storm or baseline status depending on magnitude of change
storm1 <- temp %>% 
  mutate(status = case_when(abs(diff) > daily_discharge/10 ~ "storm", abs(diff) <= daily_discharge/10 ~"base"))
         

# plot for 2017
storm1 %>% 
  filter(year(Date) == 2017) %>% 
  ggplot(aes(x = Date, y = daily_discharge)) +
  geom_line() +
  geom_point(aes(x = Date, y = daily_discharge, col = status)) 


```

Maybe, since this technique seems to capture the beginning of storm events, we could flag when the discharge *increases* by at least 10% and then also flag all the days after until the discharge returns to the level it was on the initial flagged day

```{r}
# new version of storm dataframe which flags just positive increases in discharge by over 10%
# to capture the start of a storm
storm2 <- temp %>% 
  mutate(status = case_when(diff > daily_discharge/10 ~ "storm", diff <= daily_discharge/10 ~"base"))

# plot for 2016
storm2 %>% 
  filter(year(Date) == 2015) %>% 
  filter(month(Date) %in% c(6, 7, 8, 9)) %>% 
  ggplot(aes(x = Date, y = daily_discharge)) +
  geom_line() +
  geom_point(aes(x = Date, y = daily_discharge, col = status))  


```

Below are some examples of concentration-discharge relationships during storm events.

```{r}
WQ_discharge %>% 
  filter(year(date) == 2016)  %>%
  filter((day(date) > 12 & month(date) == 7) | (day(date) < 11 & month(date) == 8)) %>% 
  ggplot(aes(x = date, y = discharge)) +
  geom_point() +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2016)  %>%
  filter((day(date) > 12 & month(date) == 7) | (day(date) < 11 & month(date) == 8)) %>% 
  ggplot(aes(x = date, y = NO3_mgL, col = site)) +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2016)  %>%
  filter((day(date) > 12 & month(date) == 7) | (day(date) < 11 & month(date) == 8)) %>% 
  arrange(date) %>% 
  ggplot(aes(x = discharge, y = NO3_mgL, col = site)) +
  geom_path() +
  facet_grid(rows = vars(site), cols = NULL, scales = "free")
  
```

```{r}
WQ_discharge %>% 
  filter(year(date) == 2017)  %>%
  filter((day(date) > 18 & day(date) < 26 & month(date) == 7) ) %>% 
  ggplot(aes(x = date, y = discharge)) +
  geom_point() +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2017)  %>%
  filter((day(date) > 18 & day(date) < 26 & month(date) == 7) ) %>% 
  ggplot(aes(x = date, y = CHLugL, col = site)) +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2017)  %>%
  filter((day(date) > 18 & day(date) < 26 & month(date) == 7) ) %>% 
  arrange(date) %>% 
  ggplot(aes(x = discharge, y = CHLugL, col = site)) +
  geom_path() 
```

```{r}
WQ_discharge %>% 
  filter(year(date) == 2018 & month(date) %in% c(6, 7))  %>%
  ggplot(aes(x = date, y = discharge)) +
  geom_point() +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2018 & month(date) %in% c(6, 7))  %>%
  ggplot(aes(x = date, y = NO3_mgL, col = site)) +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2018 & month(date) %in% c(6, 7))  %>%
  arrange(date) %>% 
  ggplot(aes(x = discharge, y = CHLugL, col = site)) +
  geom_path() 
```

```{r}
WQ_discharge %>% 
  filter(year(date) == 2015)  %>%
  filter((day(date) > 5 & month(date) == 7) ) %>% 
  ggplot(aes(x = date, y = discharge)) +
  geom_point() +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2015)  %>%
  filter((day(date) > 5 & month(date) == 7) ) %>%
  ggplot(aes(x = date, y = NO3_mgL, col = site)) +
  geom_line()

WQ_discharge %>% 
  filter(year(date) == 2015)  %>%
  filter((day(date) > 5 & month(date) == 7) ) %>%
  arrange(date) %>% 
  ggplot(aes(x = discharge, y = NO3_mgL, col = site)) +
  geom_path() 
```

