---
title: "Storm stats"
author: "Taryn Waite"
date: "6/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
WQ_hourly_discharge <- read.csv("WQ_hourly_discharge.csv")
```

```{r precip data}
precip <- read_csv("data/precip.csv") %>% 
  rename(inches = `Precip (in)`) %>% 
  mutate(inches = as.numeric(inches), Date = mdy(Date))

dailyPrecip <- precip %>% 
  drop_na() %>% group_by(Date) %>% 
  summarize(inches = sum(inches))

# function to return the cumulative precip
# during a given stretch of time given a start and end date
totalPrecip <- function(startDate, endDate){
  total <- dailyPrecip %>% 
    filter(Date >= ymd(startDate), Date <= ymd(endDate)) %>% 
    summarize(sum = sum(inches))
  return(total$sum)
}
```


This document deals with calculating various storm statistics/characteristics. Given data for the duration of a storm, the goal is to be able to produce information such as storm duration, maximum discharge, change in discharge, antecedent conditions, etc. 

First, we'll calculate storm duration. We might need to know both the length of the entire storm (given by the start and end dates) and the length of the "loop" portion of the storm (where there are matching discharge values for the rising and falling limbs, which is the portion for which we calculate HI). Here are functions for both of these:
```{r storm length stats}
# given data from the duration of the storm,
# returns the length of the storm in days
storm.length <- function(data, startDate, endDate){
  return(as.numeric(ymd(endDate)-ymd(startDate)))
}

# returns the length of the portion of the storm
# for which HI could be calculated (overlapping limbs)
storm.length.loop <- function(data){
  # find minimum discharge rows for rising and falling limbs
  mins <- data %>% filter(limb != "peak") %>% group_by(limb) %>% 
    summarise(min = min(hourlyDischarge))
  # the larger of the two is where the intervals start
  minDisLoop <- max(mins$min)
  # subset data to just the loop
  loop <- data %>% filter(hourlyDischarge >= minDisLoop)
  # get start and end dates for the loop
  start <- loop %>% summarize(start = min(dateTime))
  end <- loop %>% summarize(end = max(dateTime))
  return(as.numeric(end-start))
}
```

Next, we need a function to calculate the total change in discharge from the start of the storm to the peak. Here are 2 versions of this function. The first simply returns the change in discharge from the start of the storm (the earliest date in the storm time series) to the storm peak. The second returns the change in discharge during the loop section of the event (as described above).
**NOTE:** both of these functions use the raw data, not the smoothed data. We may want to change this later -- I'm not sure which one it makes more sense to use.
```{r discharge stats}
# returns the change in discharge (raw data)
# from the start of the event to the peak
storm.dis.change <- function(data){
  start <- (data %>% slice_min(dateTime))[["hourlyDischarge"]]
  peak <- (data %>% slice_max(hourlyDischarge))[["hourlyDischarge"]]
  return(as.numeric(peak)-as.numeric(start))
}

# returns the change in discharge (raw data)
# from the start of the overlapping rise/fall discharge
# (where the HI begins to be calculated) to the peak
storm.dis.change.loop <- function(data){
  # find minimum discharge values for rising and falling limbs
  mins <- data %>% filter(limb != "peak") %>% group_by(limb) %>% 
    summarise(min = min(hourlyDischarge))
  # the larger of the two is where the intervals start
  start <- max(mins$min)
  peak <- (data %>% slice_max(hourlyDischarge))[["hourlyDischarge"]]
  return(as.numeric(peak) - as.numeric(start))
}

# returns the maximum discharge value from the event
storm.max.dis <- function(data){
  maxDis <- data %>% summarise(maxDis = max(hourlyDischarge))
  return(as.numeric(maxDis))
}
```

It might also be interesting to see if temperature impacts c-q relationships or seasonal patterns. Here is a function to calculate the average water temperature during the event:
```{r temperature stats}
storm.avg.temp <- function(data){
  return(as.numeric(data %>% summarise(mean = mean(Temp))))
}
```

Now that we have ways of calculating event statistics, as well as HI and slope for the different constituents, we'll want to have all of the events in one dataframe with their start dates, end dates, storm statistics, and HI/slope values.

```{r function for adding events}
events <- tibble(startDate = Date(), endDate = Date(), 
                 length = numeric(), length_loop = numeric(),
                 dis_change = numeric(), dis_change_loop = numeric(),
                 max_dis = numeric(), avg_temp = numeric(),
                 HI_turb = numeric(), Slope_turb = numeric(),
                 HI_n03 = numeric(), Slope_n03 = numeric(),
                 HI_chl = numeric(), Slope_chl = numeric(),
                 HI_fdom = numeric(), Slope_fdom = numeric(),
                 HI_bga = numeric(), Slope_bga = numeric())

# adds a high-discharge event to the events tibble,
# given the data, start date, end date, and number of
# observations per day
add.event <- function(data, start, end, freq){
  # subset the data to include just the event
  subset <- data %>% 
    filter(dateTime >= ymd(start) & dateTime <= ymd(end)) %>% 
    filter(site == "MC")
  # add rising or falling limb column
  peakDate <- (subset %>% slice_max(hourlyDischarge))[["dateTime"]]
  # RL if before peakDate, FL if after, peak if on peakDate
  subset <- subset %>% 
    mutate(limb = case_when(dateTime < peakDate ~ "RL",
                            dateTime > peakDate ~ "FL",
                            T ~ "peak"))
  turb <- storm_cq(data, start, end, "Turb", 20, freq)
  n03 <- storm_cq(data, start, end, "NO3_mgL", 20, freq)
  chl <- storm_cq(data, start, end, "CHLugL", 20, freq)
  fdom <- storm_cq(data, start, end, "FDOMqsu", 20, freq)
  bga <- storm_cq(data, start, end, "BGAugL", 20, freq)
  return(events %>% add_row(startDate = ymd(start), endDate = ymd(end),
                     length = as.numeric(ymd(end) - ymd(start)), 
                     length_loop = storm.length.loop(subset),
                     dis_change = storm.dis.change(subset),
                     dis_change_loop = storm.dis.change.loop(subset),
                     max_dis = storm.max.dis(subset),
                     avg_temp = storm.avg.temp(subset),
                     HI_turb = turb[[2]], Slope_turb = turb[[3]],
                     HI_n03 = n03[[2]], Slope_n03 = n03[[3]],
                     HI_chl = chl[[2]], Slope_chl = chl[[3]],
                     HI_fdom = fdom[[2]], Slope_fdom = fdom[[3]],
                     HI_bga = bga[[2]], Slope_bga = bga[[3]]))
}
```

```{r adding events}
# initialize the tibble
events <- tibble(startDate = Date(), endDate = Date(), 
                 length = numeric(), length_loop = numeric(),
                 dis_change = numeric(), dis_change_loop = numeric(),
                 max_dis = numeric(), avg_temp = numeric(),
                 HI_turb = numeric(), Slope_turb = numeric(),
                 HI_n03 = numeric(), Slope_n03 = numeric(),
                 HI_chl = numeric(), Slope_chl = numeric(),
                 HI_fdom = numeric(), Slope_fdom = numeric(),
                 HI_bga = numeric(), Slope_bga = numeric())

# add events -- this takes a long time to run 
events <- add.event(WQ_hourly_discharge, "2016-04-25", "2016-05-24", 12)
events <- add.event(WQ_hourly_discharge, "2016-07-13", "2016-08-10", 12)
events <- add.event(WQ_hourly_discharge, "2016-06-01", "2016-06-14", 12)
events <- add.event(WQ_hourly_discharge, "2016-08-10", "2016-09-05", 12)
events <- add.event(WQ_hourly_discharge, "2016-09-06", "2016-09-20", 12)
events <- add.event(WQ_hourly_discharge, "2016-09-20", "2016-10-15", 12)
events <- add.event(WQ_hourly_discharge, "2017-06-14", "2017-06-28", 12)
events <- add.event(WQ_hourly_discharge, "2017-07-19", "2017-07-25", 24)
events <- add.event(WQ_hourly_discharge, "2017-07-26", "2017-07-30", 24)
events <- add.event(WQ_hourly_discharge, "2017-08-06", "2017-08-13", 24)
events <- add.event(WQ_hourly_discharge, "2017-09-20", "2017-09-24", 24)
events <- add.event(WQ_hourly_discharge, "2018-06-15", "2018-07-31", 12)
events <- add.event(WQ_hourly_discharge, "2015-07-06", "2015-07-28", 24)
events <- add.event(WQ_hourly_discharge, "2015-08-18", "2015-09-04", 24)
events <- add.event(WQ_hourly_discharge, "2015-09-17", "2015-09-24", 24)
events <- add.event(WQ_hourly_discharge, "2015-09-09", "2015-09-17", 24)
events <- add.event(WQ_hourly_discharge, "2017-08-16", "2017-09-17", 24)
events <- add.event(WQ_hourly_discharge, "2018-09-03", "2018-09-15", 12)


write.csv(events, "data/events.csv")
#events <- read_csv("data/events.csv")

# get rid of the duplicate rows
# (for some reason, some of the events were added multiple times)
eventsUnique <- distinct(events)

# add precip data
eventsUnique2 <- eventsUnique %>% 
  mutate(precip = mapply(totalPrecip, startDate, endDate))

```

Below, we reshape the data so that there are 4 rows for each event (one for each constituent). This makes plotting easier, so that each constituent doesn't have to be plotted separately. We also add columns for the direction of hysteresis (clockwise or counterclockwise) and slope (flushing or dilution). In Aguilera and Melack (2018), they classify clockwise hysteresis as HI >= 0.05, counterclockwise hysteresis as HI <= -0.05, flushing as $\Delta C$ >= 0.1, and dilution as $\Delta C$ <= 0.1. For now, I'm going to use 0.05 as the limit for flushing and dilution.

```{r reshaping the data}
eventsReshaped <- eventsUnique2 %>% 
  gather(key, value, -c(startDate, endDate, length, length_loop,
                        dis_change, dis_change_loop, max_dis, avg_temp, precip)) %>% 
  separate(key, c("measure", "var"), "_") %>% 
  spread(measure, value)

# add columns for directions of hysteresis and slope
eventsFinal <- eventsReshaped %>% 
  mutate(H_dir = case_when(HI >= 0.05 ~ "clockwise",
                           HI <= -0.05 ~ "counter-clockwise", T ~ "none"),
         slope_dir = case_when(Slope >= 0.05 ~ "flushing",
                               Slope <= -0.05 ~ "dilution", T ~ "constant"))
# one event (starting 9/9/15) had a value >1 for fdom slope because of negative
# values at the beginning-- might be best to take out the whole fdom row for
# this event because the data jumps from negatives to 100 (the issue we've been 
# looking at for the time series)
eventsFinal2 <- eventsFinal[-c(70,87,88),]

```

## Data exploration

```{r summarizing events}
eventsTurb <- eventsFinal2 %>% filter(var == "turb")

eventsTurb %>% 
  ggplot(aes(x = precip)) + 
  geom_histogram(bins = 5) + theme_bw()


eventsTurb %>% 
  ggplot(aes(x = max_dis)) +
  geom_histogram(bins = 5) + theme_bw()

eventsTurb %>% 
  ggplot(aes(x = dis_change)) +
  geom_histogram(bins = 5) + theme_bw()

eventsTurb %>% 
  ggplot(aes(x = avg_temp)) +
  geom_histogram(bins = 5) + theme_bw()

# base R plots look better for histograms with few events
png(filename = "event_hists.png", width = 5, height = 5,
    units = "in", res = 300, bg = "transparent")
par(mfrow = c(2,2), bg = "transparent", mgp=c(2.5,1,0), mar = c(5,2,1,2))
hist(eventsTurb$precip, breaks = 5,
     xlab = "Cumulative precipitation (in)", 
     ylab = "Number of events", main = "")
hist(eventsTurb$length, breaks = 5,
     xlab = "Length (days)", 
     ylab = "Number of events", main = "")
hist(eventsTurb$avg_temp, breaks = c(14, 17, 20, 23, 26, 29),
     xlab = "Average temperature (C)", xlim = c(10, 30),
     ylab = "Number of events", main = "")
hist(eventsTurb$dis_change, breaks = c(0, 16000, 32000, 48000, 64000, 80000),
     xlab = "Change in discharge (ft^3/s)", 
     ylab = "Number of events", main = "")
dev.off()


eventsTurb2 <- eventsTurb %>% 
  mutate(length_bin = case_when(length < 10 ~ "< 10",
                                length < 20 ~ "10-20",
                                length < 30 ~ "20-30",
                                T ~ ">= 30"),
         month = case_when(month(startDate) == 4 ~ "April",
                           month(startDate) == 6 ~ "June",
                           month(startDate) == 7 ~ "July",
                           month(startDate) == 8 ~ "August",
                           month(startDate) == 9 ~ "September"))
eventsTurb2$length_bin <- factor(eventsTurb2$length_bin, 
                                 levels = c("< 10", "10-20", "20-30", ">= 30"))
eventsTurb2$month <- factor(eventsTurb2$month,
                            levels = c("April", "June", "July", "August", "September"))

### monthly event count plot ###
png(filename = "monthly_event_count.png", width = 8, height = 6,
    units = "in", res = 300, bg = "transparent")
eventsTurb2 %>% 
  ggplot(aes(x = month, fill = as.factor(year(startDate)))) +
  geom_bar() + theme_bw() +
  theme(plot.background = element_rect(fill = "transparent",colour = NA),
         legend.background = element_rect(fill = "transparent")) +
  scale_fill_brewer(palette = "Set2", name = "Year") +
  xlab("Month of event start date") + ylab("Number of events") 
dev.off()

### version 2: heatmap ###
eventsCount <- (eventsTurb2 %>% 
  mutate(year = as.factor(year(startDate))))[,c(1, 16,17)] %>% 
  group_by(year, month) %>% tally() 

allMonthYear <- eventsCount %>% expand(month, year)

eventsCount2 <- full_join(eventsCount, allMonthYear) %>% 
  mutate(n = case_when(!is.na(n) ~ as.character(n),
                       T ~ "0"))

png(filename = "monthly_event_count_grid.png", width = 8, height = 6,
    units = "in", res = 300, bg = "transparent")
eventsCount2 %>% 
  ggplot(aes(x = year, y = month, fill = n)) +
  geom_tile(color = "black") +
  scale_fill_manual(values = c("white", "lightblue", "darkblue"),
                    name = "Number of Events") +
  theme_bw() + xlab("Year") + ylab("Month")
  theme(plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill = "transparent")) 
dev.off()  



### plot of temp over time ###
png(filename = "temp_time.png", width = 6, height = 4,
    units = "in", res = 300, bg = "transparent")
eventsTurb %>% 
  ggplot(aes(x=as.Date(paste(2014,strftime(startDate,format="%m-%d"),sep="-")), 
             y=avg_temp, 
             color=strftime(startDate,format="%Y"))) +
  geom_point(cex = 2) + theme_bw() + scale_color_brewer(palette = "Set2", name = "Year") +
  theme(plot.background = element_rect(fill = "transparent",colour = NA),
        legend.background = element_rect(fill = "transparent")) +
  xlab("Start date") + ylab("Average temperature (C)")
dev.off()

eventsTurb %>% 
  ggplot(aes(x=as.Date(paste(2014,strftime(startDate,format="%m-%d"),sep="-")), 
             y=length, 
             color=strftime(startDate,format="%Y")), size=5) +
  geom_point() + theme_bw() + scale_color_brewer(palette = "Set2", name = "Year") +
  theme(plot.background = element_rect(fill = "transparent",colour = NA))

plot(x = eventsTurb$length, y = eventsTurb$precip)

length_maxDis <- eventsTurb %>% 
  ggplot(aes(x = length, y = max_dis)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  xlab("Length (days)") + ylab("Maximum discharge (ft^3/s)") +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))

length_changeDis <- eventsTurb %>% 
  ggplot(aes(x = length, y = dis_change)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  xlab("Length (days)") + ylab("Change in discharge (ft^3/s)") +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))

png(filename = "length_dis.png", width = 6, height = 6,
    units = "in", res = 300, bg = "transparent")
ggarrange(length_changeDis, length_maxDis)
dev.off()
  
```


```{r plotting HI and slope}
# scatter plot of HI (x-axis) and slope (y-axis) for each storm
# with each constituent plotted separately

# New facet label names for constituents
var.labs <- c("Blue-green algae", "Chlorophyll", "FDOM", "Nitrate", "Turbidity")
names(var.labs) <- c("bga", "chl", "fdom", "n03", "turb")

png(filename = "HI_slope_const.png", 
    width = 9, height = 6, units = "in", res = 300, bg = "transparent")
eventsFinal2 %>% 
  ggplot(aes(x = HI, y = Slope)) +
  geom_hline(yintercept=0, linetype="dashed", color = "red") +
  geom_vline(xintercept=0, linetype="dashed", color = "red") +
  geom_point() + xlim(-1, 1) + ylim(-1, 1) +
  facet_wrap(~var, labeller = labeller(var = var.labs), nrow = 2) +
  theme_bw() + xlab("Hysteresis Index") +
  theme(plot.background = element_rect(fill = "transparent",colour = NA))
dev.off()

# looking at each storm individually
eventsFinal2 %>% 
  ggplot(aes(x = HI, y = Slope, col = var)) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  geom_vline(xintercept=0, linetype="dashed", color = "black") +
  geom_point() + xlim(-1, 1) + ylim(-1, 1) +
  facet_wrap(~as.factor(startDate)) +
  scale_color_brewer(palette = "Dark2", name = "Constituent", labels = var.labs) 

# violin plots of HI and slope for the constituents
eventsFinal2 %>% 
  ggplot(aes(x = var, y = HI)) +
  geom_violin() + ylim(-1, 1) +
  geom_hline(yintercept = 0, linetype="dashed", color = "red") +
  scale_x_discrete(labels = var.labs) + xlab("Constituent")

eventsFinal2 %>% 
  ggplot(aes(x = var, y = Slope)) +
  geom_violin() + ylim(-1, 1) +
  geom_hline(yintercept = 0, linetype="dashed", color = "red") +
  scale_x_discrete(labels = var.labs) + xlab("Constituent")

# bar plots of hysteresis direction and slope for each constituent
hiBar <- eventsFinal2 %>% 
  ggplot(aes(x = var, fill = H_dir)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_fill_brewer(palette = "Paired", name = "Hysteresis Type") +
  ylab("Number of Events") + xlab("Constituent") +
  scale_x_discrete(labels = var.labs) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  theme_bw() + 
  theme(plot.background = element_rect(fill = "transparent", colour = NA),
        legend.background = element_rect(fill = "transparent")) 
  

# need to reorder the slope directions
eventsFinal2$slope_dir <- factor(eventsFinal2$slope_dir, 
                                 levels = c("flushing", "dilution", "constant"))
slopeBar <- eventsFinal2 %>% 
  ggplot(aes(x = var, fill = slope_dir)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Paired", name = "Slope Type") +
  ylab("Number of Events") + xlab("Constituent") +
  scale_x_discrete(labels = var.labs) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) 

png(filename = "HI_slope_bar.png", 
    width = 8, height = 8, units = "in", res = 300, bg = "transparent")
ggarrange(hiBar, slopeBar, nrow = 2)
dev.off()
```

```{r event stats vs hi/slope}
# scatterplots showing relationships between HI/slope and different
# discharge event characteristics
eventsFinal2 %>% 
  ggplot(aes(x = dis_change, y = HI)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = dis_change, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = length, y = HI)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = length, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = avg_temp, y = HI)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = avg_temp, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal3 <- eventsFinal2 %>% 
  mutate(avg_rate_dis = dis_change/length)

eventsFinal3 %>% 
  ggplot(aes(x = avg_rate_dis, y = HI)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal3 %>% 
  ggplot(aes(x = avg_rate_dis, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = precip, y = HI)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)

eventsFinal2 %>% 
  ggplot(aes(x = precip, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~var)
```

```{r}
eventsN03 <- eventsFinal2 %>% filter(var == "n03")
eventsChl <- eventsFinal2 %>% filter(var == "chl")
eventsBGA <- eventsFinal2 %>% filter(var == "bga")
eventsFDOM <- eventsFinal2 %>% filter(var == "fdom")

png(filename = "dis_change_hi_allVar.png", width = 8, height = 6, 
    units = "in", res = 300, bg = "transparent")
eventsFinal2 %>% 
  ggplot(aes(x = dis_change, y = HI, col = var)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  scale_color_manual(values = myColors, name = "Constituent", labels = var.labs) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent"))
dev.off()

lm_disChange_turb_HI <- lm(eventsTurb$HI ~ eventsTurb$dis_change)
summary(lm_disChange_turb_HI)
lm_disChange_n03_HI <- lm(eventsN03$HI ~ eventsN03$dis_change)
summary(lm_disChange_n03_HI)
lm_disChange_chl_HI <- lm(eventsChl$HI ~ eventsChl$dis_change)
summary(lm_disChange_chl_HI)
lm_disChange_bga_HI <- lm(eventsBGA$HI ~ eventsBGA$dis_change)
summary(lm_disChange_bga_HI)
lm_disChange_fdom_HI <- lm(eventsFDOM$HI ~ eventsFDOM$dis_change)
summary(lm_disChange_fdom_HI)

lm_disChange_turb_slope <- lm(eventsTurb$Slope ~ eventsTurb$dis_change)
summary(lm_disChange_turb_slope)
lm_disChange_n03_slope <- lm(eventsN03$Slope ~ eventsN03$dis_change)
summary(lm_disChange_n03_slope)
lm_disChange_chl_slope <- lm(eventsChl$Slope ~ eventsChl$dis_change)
summary(lm_disChange_chl_slope)
lm_disChange_bga_slope <- lm(eventsBGA$Slope ~ eventsBGA$dis_change)
summary(lm_disChange_bga_slope)
lm_disChange_fdom_slope <- lm(eventsFDOM$Slope ~ eventsFDOM$dis_change)
summary(lm_disChange_fdom_slope)

# sig: dis_change/HI (FDOM and turb); avg_temp/Slope (turb); avg_temp/HI (nO3)

# plots of significant (ish) linear models 
# dis change vs hi (fdom)
eventsFDOM %>% 
  ggplot(aes(x = dis_change, y = HI)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  xlab("Change in discharge (ft^3/s)") + ylab("HI") +
  annotate(geom="text", x=60000, y=0.75, 
           label="Dissolved organic matter", col = "blue", cex = 5)

# dis change vs hi (turb)
eventsTurb %>% 
  ggplot(aes(x = dis_change, y = HI)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  xlab("Change in discharge (ft^3/s)") + ylab("HI") +
  annotate(geom="text", x=60000, y=0.75, 
           label="Turbidity", col = "blue", cex = 5)

## both fdom and turb on same plot for dis change vs hi
eventsFDOM_turb <- eventsFinal2 %>% 
  filter(var %in% c("turb", "fdom"))

png(filename = "dis_change_hi_turb_fdom.png", width = 6, height = 4,
    units = "in", res = 300, bg = "transparent")
eventsFDOM_turb %>% 
  ggplot(aes(x = dis_change, y = HI, col = var)) +
  geom_point() + geom_smooth(method = "lm", se = F) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  xlab("Change in discharge (ft^3/s)") + ylab("Hysteresis Index") +
  scale_color_manual(values = myColors, labels = var.labs, name = "Constituent") +
  annotate(geom = "text", x = 50000, y = 0.8, hjust = 0, 
           label = "p = 0.040", col = myColors[3]) +
  annotate(geom = "text", x = 50000, y = 0.7, hjust = 0,
           label = "R-squared = 0.24", col = myColors[3]) +
  annotate(geom = "text", x = 40000, y = -0.55, hjust = 0,
           label = "p = 0.035", col = myColors[2]) +
  annotate(geom = "text", x = 40000, y = -0.65, hjust = 0,
           label = "R-squared = 0.26", col = myColors[2])
dev.off()

# avg temp vs HI (n03)
png(filename = "temp_hi_nO3.png", width = 6, height = 4,
    units = "in", res = 300, bg = "transparent")
eventsN03 %>% 
  ggplot(aes(x = avg_temp, y = HI)) +
  geom_point(col = myColors[4]) + 
  geom_smooth(method = "lm", se = F, col = myColors[4]) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  xlab("Average temperature (C)") + ylab("Hysteresis Index") +
  annotate(geom = "text", x = 17, y = -0.1, hjust = 0,
           label = "p = 0.063", col = myColors[4], cex = 4) +
  annotate(geom = "text", x = 17, y = -0.2, hjust = 0,
           label = "R-squared = 0.20", col = myColors[4], cex = 4)
dev.off()

# avg temp vs slope (turb)
png(filename = "temp_slope_turb.png", width = 6, height = 4,
    units = "in", res = 300, bg = "transparent")
eventsTurb %>% 
  ggplot(aes(x = avg_temp, y = Slope)) +
  geom_point(col = myColors[3]) + 
  geom_smooth(method = "lm", se = F, col = myColors[3]) +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  xlab("Average temperature (C)") + ylab("Slope") +
  annotate(geom = "text", x = 17.5, y = 0.23, hjust = 0,
           label = "p = 0.054", col = myColors[3]) +
  annotate(geom = "text", x = 17.5, y = 0.19, hjust = 0,
           label = "R-squared = 0.21", col = myColors[3])
dev.off()



```


```{r pairwise constituent comparison}
var.labs <- c("Blue-green algae", "Chlorophyll", "FDOM", "Nitrate", "Turbidity")
names(var.labs) <- c("bga", "chl", "fdom", "n03", "turb")

# create a color palette that maps to the variable names
myColors <- brewer.pal(5,"Dark2")
names(myColors) <- c("bga", "fdom", "turb", "n03", "chl")

# create a list of all possible variable pairs
pairs <- list(c("bga", "chl"), c("bga", "fdom"), c("bga", "n03"), 
              c("bga", "turb"), c("chl", "fdom"), c("chl", "n03"),
              c("chl", "turb"), c("fdom", "n03"), c("fdom", "turb"), 
              c("n03", "turb"))

# make a list of plots of the variable pairs with lines connecting
# corresponding points (same storm, different constituent)
plots <- list()
for(i in 1:10){
  pair <- pairs[[i]]
  p <- ggplot(subset( eventsFinal2, var %in% pair), 
         aes(x = HI, y = Slope, group = as.factor(startDate), col = var)) +
        geom_hline(yintercept=0, linetype="dashed", color = "black") +
        geom_vline(xintercept=0, linetype="dashed", color = "black") +
        geom_line(col = "grey") +
        geom_point() + xlim(c(-1,1)) + ylim(c(-1, 1)) + xlab("Hysteresis Index") +
        scale_color_manual(name = "Constituent", values = myColors, labels = var.labs) +
        theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) 
  plots[[i]] <- p
}

#ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]])
#ggarrange(plots[[1]], plots[[5]], plots[[6]], plots[[7]])
#ggarrange(plots[[2]], plots[[5]], plots[[7]], plots[[8]])
#ggarrange(plots[[4]], plots[[7]], plots[[9]], plots[[10]])
#ggarrange(plots[[3]], plots[[6]], plots[[8]], plots[[10]])

png(filename = "bga-chl.png", width = 6, height = 4, 
    res = 300, units = "in", bg = "transparent")
plots[[1]]
dev.off()

plots[[2]]
plots[[3]]
plots[[4]]
plots[[5]]
plots[[6]]

png(filename = "chl-turb.png", width = 6, height = 4, 
    res = 300, units = "in", bg = "transparent")
plots[[7]]
dev.off()

plots[[8]]
plots[[9]]
plots[[10]]
```

After looking at the pairwise comparison plots, it might also be interesting to compute a distance matrix between constituents for each storm.

```{r distance matrix}
# function which, given 2 sets of HI and slope values,
# returns the Euclidean distance between the 2
dist <- function(HI_1, HI_2, slope_1, slope_2){
  return(sqrt((HI_1-HI_2)^2 + (slope_1-slope_2)^2))
}

# remove events where we don't have HI and slope for all constituents
# (may want to figure out a way to put these back in and use 
# just the constituents that are available)
eventsUnique4 <- eventsUnique2[-c(15, 16),]

# initialize the distance matrix
dists <- matrix(nrow = 5, ncol = 5, data = 0)
# list of the constituents
vars <- c("bga", "chl", "fdom", "n03", "turb")
# for each event, add the appropriate distance to each
# slot in the matrix
for(i in 1:nrow(eventsUnique4)){
  for(j in 1:5){
    for(k in 1:5){
      dists[j,k] <- dists[j,k] + 
        dist(eventsUnique4[[i,paste0("HI_", vars[j])]], 
             eventsUnique4[[i,paste0("HI_", vars[k])]],
             eventsUnique4[[i,paste0("Slope_", vars[j])]],
             eventsUnique4[[i,paste0("Slope_", vars[k])]])
    }
  }
}
# divide by the number of events to obtain average distances
dists <- dists/nrow(eventsUnique4)

rownames(dists) <- c("BGA", "CHL", "FDOM", "NO3", "TURB")
colnames(dists) <- c("BGA", "CHL", "FDOM", "NO3", "TURB")



# data wrangling in preparation for heatmap plotting
#dists_t <- t(dists)
dists[upper.tri(dists, diag = T)] <- NA

dists_df <- dists %>%
  as_tibble() %>%
  rowid_to_column(var="X") %>%
  gather(key="Y", value="Distance", -1) %>% 
  mutate(X = case_when(X == 1 ~ "BGA",
                       X == 2 ~ "CHL",
                       X == 3 ~ "FDOM",
                       X == 4 ~ "NO3",
                       X == 5 ~ "TURB"))



png(filename = "dist_heatmap_vir.png", width = 8, height = 6,
    unit = "in", res = 300, bg = "transparent")
# heatmap of the distance matrix
ggplot(dists_df, aes(x = X, y = Y, fill = Distance)) + 
  geom_tile() +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  #scale_fill_continuous(na.value = 'white') +
  scale_fill_viridis_c(na.value = "white") +
  xlab("") + ylab("") +
  scale_x_discrete(limits = rev(levels(as.factor(dists_df$X)))) 
dev.off()
  
```


Clustering with just hysteresis index and slope:
```{r clusters with hi and slope}
turbDat <- eventsFinal2 %>% filter(var == "turb")
hi_slope_dist_turb <- dist(turbDat[, c(10,11)])
hi_slope_complete_turb <- hclust(hi_slope_dist_turb, method = "complete")
plot(hi_slope_complete_turb, labels = turbDat$startDate)
hi_slope_turb_3 <- cutree(hi_slope_complete_turb, k = 3)

plot(x = turbDat$HI, y = turbDat$Slope, col = hi_slope_turb_3, pch = 20)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")


n03Dat <- eventsFinal2 %>% filter(var == "n03")
hi_slope_dist_n03 <- dist(n03Dat[, c(10,11)])
hi_slope_complete_n03 <- hclust(hi_slope_dist_n03, method = "single")
plot(hi_slope_complete_n03, labels = n03Dat$startDate)
hi_slope_n03_2 <- cutree(hi_slope_complete_n03, k = 2)

plot(x = n03Dat$HI, y = n03Dat$Slope, col = hi_slope_n03_2, pch = 20)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")


chlDat <- eventsFinal2 %>% filter(var == "chl")
hi_slope_dist_chl <- dist(chlDat[, c(10,11)])
hi_slope_complete_chl <- hclust(hi_slope_dist_chl, method = "single")
plot(hi_slope_complete_chl, labels = chlDat$startDate)
hi_slope_chl_3 <- cutree(hi_slope_complete_chl, k = 3)

plot(x = chlDat$HI, y = chlDat$Slope, col = hi_slope_chl_3, pch = 20)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")


fdomDat <- eventsFinal2 %>% filter(var == "fdom")
hi_slope_dist_fdom <- dist(fdomDat[, c(10,11)])
hi_slope_complete_fdom <- hclust(hi_slope_dist_fdom, method = "single")
plot(hi_slope_complete_fdom, labels = fdomDat$startDate)
hi_slope_fdom_4 <- cutree(hi_slope_complete_fdom, k = 4)

plot(x = fdomDat$HI, y = fdomDat$Slope, col = hi_slope_fdom_4, pch = 20)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")

bgaDat <- eventsFinal2 %>% filter(var == "bga")
hi_slope_dist_bga <- dist(bgaDat[, c(10,11)])
hi_slope_complete_bga <- hclust(hi_slope_dist_bga, method = "single")
plot(hi_slope_complete_bga, labels = bgaDat$startDate)
hi_slope_bga_2 <- cutree(hi_slope_complete_bga, k = 2)

plot(x = bgaDat$HI, y = bgaDat$Slope, col = hi_slope_bga_2, pch = 20)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")
```

```{r clusters with hi/slope for all variables}
hi_slope_dist <- dist(eventsFinal2[, c(10,11)])
hi_slope_complete <- hclust(hi_slope_dist, method = "complete")
plot(hi_slope_complete, labels = eventsFinal2$startDate)
hi_slope_5 <- cutree(hi_slope_complete, k = 5)

plot(x = eventsFinal2$HI, y = eventsFinal2$Slope, 
     col = hi_slope_5, pch = eventsFinal2$var,)
abline( h = 0, lty = 2, col = "red") ; abline(v = 0, lty = 2, col = "red")
```


```{r clusters with all variables HI/slope}
# need to fix this because of the weird fdom event 
# (only fixed in the eventsFinal df, not the eventsUnique one)
allVar_hi_dist <- dist(eventsUnique[, c(9, 11, 13, 15)])
allVar_hi_complete <- hclust(allVar_hi_dist, method = "complete")
plot(allVar_hi_complete, labels = eventsUnique$startDate)

allVar_hi_3 <- cutree(allVar_hi_complete, k = 3)
pairs(eventsUnique[, c(9, 11, 13, 15)], panel = function(x,y) text(x, y, allVar_hi_3))

allVar_slope_dist <- dist(eventsUnique[,c(10, 12, 14, 16)])
allVar_slope_complete <- hclust(allVar_slope_dist, method = "complete")
plot(allVar_slope_complete, labels = eventsUnique$startDate)


allVar_dist <- dist(eventsUnique[, -c(1:8)])
allVar_complete <- hclust(allVar_dist, method = "complete")
plot(allVar_complete, labels = eventsUnique$startDate)

allVar_3 <- cutree(allVar_complete, k = 3)
pairs(eventsUnique[, -c(1:8)], panel = function(x,y) text(x, y, allVar_3))
```

```{r clustering events by stats}
events_dist <- dist(eventsUnique[,c(5, 7, 8)])
events_complete <- hclust(events_dist, method = "complete")
plot(events_complete, labels = eventsUnique$startDate)

events_3 <- cutree(events_complete, k = 3)
pairs(eventsUnique[, c(5, 7, 8)], panel = function(x,y) text(x, y, events_3))
```

```{r example plot for storm ID}


ggplot(discharge_2017, aes(x = date, y = discharge)) + 
  geom_line() + xlab("Time") + ylab("Discharge (ft^3/s)") +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent"))

png(filename = "stormID_2016.png", width = 6, height = 4,
    units = "in", res = 300, bg = "transparent")
ggplot(discharge_2016, aes(x = date, y = discharge)) + 
  geom_line() + xlab("Time") + ylab("Discharge (ft^3/s)") +
  theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA),
                     legend.background = element_rect(fill = "transparent")) +
  geom_point(data = tibble(X = c(ymd("2016-04-25"), ymd("2016-07-13"), ymd("2016-06-01"), 
                   ymd("2016-08-10"),ymd("2016-09-06"), ymd("2016-09-20")),
              Y = c(34500, 32500, 46000, 36700, 43900, 50800)), aes(x = X, y = Y), col = "red", cex = 2)
dev.off()
  
```


