---
title: "Week 4"
author: "James Pack"
date: "6/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggthemes)
library(dplyr)
```

```{r data prep, include = F}
phytos_9606 <- read_csv('Phytos_1996-2006.csv')
phytos_0812 <- read_csv('Phytos_2008-2012.csv')
phyto_info <- read_csv('Phyto_info_WQdata.csv')

# Combining data from all time periods
phytos_whole <- phytos_9606 %>% bind_rows(phytos_0812)

# Adding water quality data, joined by SHEETBAR (ID) of sample
phyto <- phytos_whole %>% left_join(phyto_info,by="SHEETBAR")

# Cleaning the dates
phyto <- phyto %>% mutate(DATE=mdy(DATE))
```

## Community Metrics

Adding in measures for Total Abundance, Richness, Relative Cyanobacteria Abundance, Cyanobacteria Richness, and Non-Cyanobacteria Richness. Need to edit these so that NAs are given 0s

```{r adding metrics}
# Richness
phyto <- left_join(phyto,phyto %>% count(SHEETBAR),by='SHEETBAR') %>% 
  rename(tot_rich=n)
# Total Abundance
total <- phyto %>% group_by(SHEETBAR) %>% 
  summarize(total_abundance = sum(`TOTAL BV_um3L`))
phyto <- left_join(phyto,total,by='SHEETBAR')
# Cyanobacteria Abundance
cyan_abundance <- phyto %>% filter(DIVISION == 'Cyanobacteria') %>% 
  group_by(SHEETBAR) %>% 
  summarize(cyan_abund = sum(`TOTAL BV_um3L`))
phyto <- left_join(phyto,cyan_abundance,by='SHEETBAR')
phyto %>% mutate(case_when(is.na(cyan_abund) == T ~ 0))
# Relative Cyanobacteria Abundance
phyto <- phyto %>% mutate(rel_cyan = cyan_abund/total_abundance * 100)
# Cyanobacteria Richness
cyan_richness <- phyto %>% filter(DIVISION == 'Cyanobacteria') %>% 
  count(SHEETBAR)
phyto <- left_join(phyto,cyan_richness,by='SHEETBAR') %>% 
  rename(cyan_rich=n)
phyto %>% mutate(case_when(is.na(cyan_rich) == T ~ 0))
# Non-Cyanobacteria Abundance
phyto <- phyto %>% mutate(non_cyan_abund = total_abundance - cyan_abund)
# Relative Non-Cyanobacteria Abundance
phyto <- phyto %>% 
  mutate(rel_non_cyan_abund = non_cyan_abund/total_abundance * 100)
# Non-Cyanobacteria Richness
phyto <- phyto %>% mutate(non_cyan_rich = tot_rich - cyan_rich)
```

## Graphing New Metrics
Time to make some plots & see what's happening.

```{r graphing metrics over time}
phyto %>% ggplot(aes(DATE,total_abundance)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Abundance Over Time') +
  xlab('Date') +
  ylab('Biovolume') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,cyan_abund)) +
  geom_point() + 
  geom_smooth(method=lm) +
  ggtitle('Cyanobacteria Abundance Over Time') +
  xlab('Date') +
  ylab('Biovolume') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,non_cyan_abund)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Non-Cyanobacteria Abundance Over Time') +
  xlab('Date') +
  ylab('Biovolume') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,tot_rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Total Richness Over Time') +
  xlab('Date') +
  ylab('# of Genera/Species Observed') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,cyan_rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Cyanobacteria Richness Over Time') +
  xlab('Date') +
  ylab('# of Genera/Species Observed') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,non_cyan_rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Non-Cyanobacteria Richness Over Time') +
  xlab('Date') +
  ylab('# of Genera/Species Observed') +
  theme_gdocs()
phyto %>% ggplot(aes(DATE,cyan_rich/non_cyan_rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  ggtitle('Ratio of Cyanobacteria Richness to Rest of Community') +
  xlab('Date') +
  ylab('# of Cyanobacteria/# of other Phytoplankton') +
  theme_gdocs()
```

No clear trends that I see yet. Time to dig deeper. 

```{r graphing metrics against each other}
phyto %>% ggplot(aes(cyan_rich,non_cyan_rich)) +
  geom_point() 
phyto %>% ggplot(aes(rel_cyan,non_cyan_rich)) +
  geom_point() 
phyto %>% ggplot(aes(rel_cyan,cyan_rich)) +
  geom_point()
phyto %>% ggplot(aes(cyan_abund,non_cyan_abund)) +
  geom_point()
phyto %>% ggplot(aes(log(cyan_abund),log(non_cyan_abund))) +
  geom_point() +
  geom_smooth(method=lm)
```

## Cyanobacteria Project Document

This section is work on the document Kathi Jo created to guide the Cyanobacteria focus.

### Question 1

"What is the abundance and richness of cyanobacteria species across pools of the UMRS?  What species/genera of cyanobacteria are most common and where?"

```{r cyano question 1}
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(DATE,cyan_abund/1e9)) +
  geom_point() +
  geom_smooth(method=lm) +
  facet_wrap(~FLDNUM,nrow=3)
phyto %>% filter(is.na(FLDNUM) == F) %>% 
  ggplot(aes(DATE,cyan_rich)) +
  geom_point() +
  geom_smooth(method=lm) +
  facet_wrap(~FLDNUM,nrow=3)
```

There seems to have been more cyanobacteria at field stations 1 and 2 more than the other stations. Station 3 had one large outlier around 2008 that will be interesting to look into, but otherwise is like stations 4, 5, and 6 with generally lower values. Stations 1 and 2 also had richer cyanobacteria communities, while the other four stations tended to have fewer varieties. Now to figure out what those cyanobacteria genera/species were.

```{r cyano species}
cyan_at_pool <- function(pool_num){
  return(unique((phyto %>% filter(DIVISION == 'Cyanobacteria' & FLDNUM == pool_num))$GENUS))
}
cyan_pool_1 <- cyan_at_pool(1)
cyan_pool_2 <- cyan_at_pool(2)
cyan_pool_3 <- cyan_at_pool(3)
cyan_pool_4 <- cyan_at_pool(4)
cyan_pool_5 <- cyan_at_pool(5)
cyan_pool_6 <- cyan_at_pool(6)
```

These six vectors now have the cyanobacteria that were at some point observed at their respective field station. Now to find which cyanobacteria were present at all of the stations.

```{r cyano species common}
present_everywhere <- intersect(intersect(intersect(cyan_pool_1,cyan_pool_2), intersect(cyan_pool_3,cyan_pool_4)), intersect(cyan_pool_5,cyan_pool_6))
print(present_everywhere)
```

The above 8 cyanobacteria were observed at all six stations, making them the most common cyanobacteria in the data. The following are the unique cyanobacteria to each field station.

```{r cyano species unique}
pool_1_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_1,cyan_pool_2), cyan_pool_3), cyan_pool_4), cyan_pool_5), cyan_pool_6)
pool_2_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_2,cyan_pool_1), cyan_pool_3), cyan_pool_4), cyan_pool_5), cyan_pool_6)
pool_3_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_3,cyan_pool_2), cyan_pool_1), cyan_pool_4), cyan_pool_5), cyan_pool_6)
pool_4_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_4,cyan_pool_2), cyan_pool_3), cyan_pool_1), cyan_pool_5), cyan_pool_6)
pool_5_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_5,cyan_pool_2), cyan_pool_3), cyan_pool_4), cyan_pool_1), cyan_pool_6)
pool_6_uniq <- setdiff(setdiff(setdiff(setdiff(setdiff(cyan_pool_6,cyan_pool_2), cyan_pool_3), cyan_pool_4), cyan_pool_5), cyan_pool_1)
```

Somewhat surprisingly, none of the six pool had unique cyanobacteria present. That is, every cyanobacteria that was observed was present in at least two of the pools. Additionally, because the highest number of unique cyanobacteria observed at a given pool was 16 and 8 were common to all stations, the past observation that more general models are stronger than river/station specific models may hold true in future work as well. 
